---
title: "Mammal data - Hearbaceus and dwarf shrub vegetation"
author: "Ella Pasternak and Talia Gabay"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)
#knitr::opts_chunk$set(dev = "cairo_pdf", dev.args=list(cairo_pdf = list(family='Arial')))

#Load packages
library(devtools)
library(rlang)
library(ggplot2)
library(data.table)
library(dplyr)
library(mvabund)
library(kableExtra)
library(lme4)
library(MASS)
library(vegan)
library(jtools)
library(interactions)
library(car)
library(ecoCopula)
library(chron)
library(readxl)
library(performance)
library(glmmTMB) 
library(emmeans)
library(Cairo)
library(extrafont)

#Load functions
source("../functions/plot_model_effect_mammals_batha.R")
source("../functions/plot_interaction_mammals_batha.R")

```


```{r load data, include=FALSE}
#Batha data - load Rdata file
load("../data/Herb_analysis_2023.RData")

#Load table for species richness
all.data.mammals.final.final.for.richness <- read.csv("../data/all.data.mammals.final.final.for.richness.csv", header = TRUE, sep =",") %>% as.data.table()

```


```{r data exploration}
#find non- regular values with abundance:
D <- Herb_abu %>% as.data.table()
D$TotAb <- rowSums(D[,1:7])
D$TotRi <- rowSums(D[,1:7]>0)
View(D)

my_cols <- c("blue", "red")
dotchart(D$TotAb, ylab = "Site",
         groups = env_data.herb$Site, 
         # color = my_cols[sp.richness.Maquis.per.cam$Settlements],
         cex = 0.9,  pch = 1, xlab = "abundance")


my_cols <- c("blue", "red","darkgreen","brown","purple")
dotchart(D$TotAb, ylab = "Campaign",gcolor = my_cols,
         groups = env_data.herb$Project.ID, 
         color = my_cols[env_data.herb$Project.ID],
         cex = 0.9,  pch = 1, xlab = "abundance")

Da <- Herb.mammals.abundance.matrix %>% as.data.table()
Da$TotAb <- rowSums(Da[,18:24])


kable(summary(Da[,.(TotAb,Project.ID, Settlements, Site,rescaled_Time.Diff_new, cosinus_Monitoring.Time.Diff, sinus_Monitoring.Time.Diff,Transect, Distance, Distance.Agri)]),"pipe")

pairs(Da[,lapply(X = .SD,FUN = as.numeric),.SDcols=c("Settlements","Project.ID","Site","rescaled_Time.Diff_new", "sinus_Monitoring.Time.Diff", "cosinus_Monitoring.Time.Diff", "Transect", "Distance","Distance.Agri")])

kable(cor(Da[,lapply(X = .SD,FUN = as.numeric),.SDcols=c("Settlements","Project.ID","Site","rescaled_Time.Diff_new", "sinus_Monitoring.Time.Diff", "cosinus_Monitoring.Time.Diff", "Transect", "Distance","Distance.Agri")]),"pipe")

```

#seems that Project.ID and rescaled.time.diff has strong correaltion, obviously.
#Site and transect has strong correlation

#the distance has some correlation with site (-0.7), and transect (-0.61/-0.68), and with distance.Agri (-0.82).
#distance.agri has correlation with site (-0.82) and transect (-0.79/-0.82) and with distance (0.75).
#it might be problematic to put site and distance together in the model, because of the high correlation.

```{r data preparation}
#Make Site as factor
env_data.herb <- env_data.herb[,Site:=factor(Site)]

#Factor levels
unique(env_data.herb$Site)

```

#### Community analysis using package MVabund


```{r Compare model distributions}
#Compare manyglm model distributions
mva_herb_poiss <- manyglm(Herb_spp~ Distance_rescaled*rescaled_Time.Diff_new+Site + 
                      sinus_Monitoring.Time.Diff+cosinus_Monitoring.Time.Diff , family = "poisson", data=env_data.herb)

plot.manyglm(mva_herb_poiss, which = 1:3)

mva_herb_nb <- manyglm(Herb_spp~ Distance_rescaled*rescaled_Time.Diff_new+Site + 
                            sinus_Monitoring.Time.Diff+cosinus_Monitoring.Time.Diff , family = "negative.binomial", data=env_data.herb)

plot.manyglm(mva_herb_nb, which = 1:3)

```

#mva_herb_nb looks much better.

```{r model selection}
#model selection:
drop1(mva_herb_nb)
```

#nothing has to be dropped. 

#the final model is mva_herb_nb

```{r final model}
anov_mva_herb_nb <- anova(mva_herb_nb, p.uni = "adjusted")
print(anov_mva_herb_nb)

#graph:
coefp_Herb <- merge(data.table(t(coef(mva_herb_nb)),keep.rownames=TRUE), data.table(t(anov_mva_herb_nb$uni.p), keep.rownames = TRUE), by = "rn") # coefficients are exponentiated because the link function used for poisson is log -> above 1 is positive and below 1 is negative

# add total species abundance
coefp_Herb <- merge(coefp_Herb,as.data.table(colSums(Herb_spp),keep.rownames = TRUE), by.x = "rn", by.y = "V1")

colnames(coefp_Herb)

colnames(coefp_Herb) <- c("SciName","Intercept.coef","Distance.coef","Time_diff.coef","Site_Karei.Deshe.coef","Site_Natur.coef","Site_Shaal.coef","Site_Yavneel.coef", "sinus_monitoring.time.coef","cosinus_monitoring.time.coef","Distance_Time.coef", "Intercept.p","Distance.p", "Time_diff.p","Site.p", "sin_time.p","Cosin_time.p","Distance_Time.p" ,"species_abundance")


#add hebrew name
A <- as.data.table(all.data.mammals.final.final.without.rare)
sci_heb <- A[,.(Species_Heb=unique(Species_Heb)), keyby=Species]
sci_heb[,Species:=make.names(Species)]
coefp_Herb <- merge(coefp_Herb, sci_heb, by.x = "SciName", by.y = "Species")

write.csv(coefp_Herb, "../output/batha/coefficients_Herb.csv")
```


#only golden jackal, grey wolf and red fox are affected by the parameters in the model.



```{r community analysis - per species models - Jackal}

####Canis aureus####

golden.jackal.herb <- Herb_abundmat[,c(1:18,25:26)]

names(golden.jackal.herb)[names(golden.jackal.herb) == 'Canis aureus'] <- 'Canis.aureus'

#only distance is significant, so I run model without intercation:


glm.golden.jackal.herb <- glm.nb(Canis.aureus ~ Distance_rescaled+rescaled_Time.Diff_new+Site + 
                                   sinus_Monitoring.Time.Diff+cosinus_Monitoring.Time.Diff, data = golden.jackal.herb, maxit = 999)


summary(glm.golden.jackal.herb)


#no dots:
effect_plot_g.Jackal.herb <- effect_plot(model = glm.golden.jackal.herb, data=golden.jackal.herb, pred = Distance_rescaled, interval = T, plot.points = F, jitter = c(0.1,0), point.size = 3, cat.pred.point.size = 4, partial.residuals = F, colors = "Qual1",line.colors = "black") 
effect_plot_g.Jackal.herb+ylim(0,200)

my_cols <- c("blue", "red")
dotchart(golden.jackal.herb$Canis.aureus, ylab = "Site",
         groups = golden.jackal.herb$Site, 
         # color = my_cols[sp.richness.Maquis.per.cam$Settlements],
         cex = 0.9,  pch = 1, xlab = "abundance")

#gcolor = my_cols,


my_cols <- c("blue", "red","darkgreen","brown","purple")
dotchart(golden.jackal.herb$Canis.aureus, ylab = "Campaign",gcolor = my_cols,
         groups = golden.jackal.herb$Project.ID, 
         color = my_cols[golden.jackal.herb$Project.ID],
         cex = 0.9,  pch = 1, xlab = "abundance")

#find how many jackals are in each site and each campaign:
jackal <- as.data.table(golden.jackal.herb)
jackal.counts <- jackal[,.(total.count = sum(Canis.aureus)), keyby=c("Project.ID","Site")]


p <- ggplot(golden.jackal.herb, aes(x=Proje, y=Canis.aureus))+ geom_point(aes(color=Site),size=5)+
  theme(axis.text=element_text(size=12),axis.title=element_text(size=14,face="bold"))

write.csv(jackal.counts, "../output/batha/jackal.counts_Herb.csv")

#there are some sites with less jackal obs, but not completely zeros. So, I do not know if there is
#justification to remove sites. 

p7 <- ggplot(golden.jackal.herb, aes(x=Project.ID, y=Canis.aureus))+ geom_point(aes(color=Site),size=5)+
  theme(axis.text=element_text(size=12),axis.title=element_text(size=14,face="bold"))
# scale_y_continuous(limits = c(0,100))

p7

p5 <- ggplot(golden.jackal.herb, aes(fill=Site,y=Canis.aureus,x=Project.ID))+ geom_bar(position="dodge", stat="identity")+
  theme(axis.text=element_text(size=12),axis.title=element_text(size=14,face="bold"))
# scale_y_continuous(limits = c(0,100))

p5
```


```{r community analysis (subunit) - per species models - wolf}

rare.species.wolf <- Rare.sp_by.two.conditions[Rare.sp_by.two.conditions$Species=="Canis lupus",]


write.csv(rare.species.wolf, "../output/batha/rare.species.wolf.csv")

#grey wolf:
wolf.herb <- Herb_abundmat[,c(1:17,19,25:26)]

unique(wolf.herb$Site)

names(wolf.herb)[names(wolf.herb) == 'Canis lupus'] <- 'Canis.lupus'

#only the interaction between distance and time is significant:


glm.wolf.herb <- glm.nb(Canis.lupus ~ Distance_rescaled*rescaled_Time.Diff_new+Site + 
                                   sinus_Monitoring.Time.Diff+cosinus_Monitoring.Time.Diff, data = wolf.herb, maxit = 999)


summary(glm.wolf.herb)

#the estimate numbers are really low.

my_cols <- c("blue", "red")
dotchart(wolf.herb$Canis.lupus, ylab = "Site",
         groups = wolf.herb$Site, 
         # color = my_cols[sp.richness.Maquis.per.cam$Settlements],
         cex = 0.9,  pch = 1, xlab = "abundance")

#gcolor = my_cols,


my_cols <- c("blue", "red","darkgreen","brown","purple")
dotchart(wolf.herb$Canis.lupus, ylab = "Campaign",gcolor = my_cols,
         groups = wolf.herb$Project.ID, 
         color = my_cols[wolf.herb$Project.ID],
         cex = 0.9,  pch = 1, xlab = "abundance")

#remove Karei Deshe:
wolf.herb_no.KD <- wolf.herb[!Site %in% "Karei Deshe",]

wolf.herb_no.KD <- wolf.herb_no.KD[,Site:=factor(Site)]

unique(wolf.herb_no.KD$Site)

glm.wolf.herb2 <- glm.nb(Canis.lupus ~ Distance_rescaled*rescaled_Time.Diff_new+Site + 
                          sinus_Monitoring.Time.Diff+cosinus_Monitoring.Time.Diff, data = wolf.herb_no.KD, maxit = 999)


summary(glm.wolf.herb2)

#the estimates are better.

wolf_interaction <- interact_plot(model = glm.wolf.herb2, data=wolf.herb_no.KD, pred = rescaled_Time.Diff_new, modx =Distance_rescaled , partial.residuals = F, jitter = c(0.2), point.size = 3, rug = F, point.shape = F, interval = T,modx.values = "plus-minus")
wolf_interaction 
#no interval:
interact_plot(model = glm.wolf.herb2, data=wolf.herb_no.KD, pred = rescaled_Time.Diff_new, modx =Distance_rescaled , partial.residuals = F, jitter = c(0.2), point.size = 3, rug = F, point.shape = F, interval = F,modx.values = "plus-minus")

#for wolf- try to remove Karei Deshe and Natur too:

wolf.herb_thin <- subset(wolf.herb, Site!="Karei Deshe" & Site!="Natur")

wolf.herb_thin <- as.data.table(wolf.herb_thin)

wolf.herb_thin <- wolf.herb_thin[,Site:=factor(Site)]

unique(wolf.herb_thin$Site)

glm.wolf.herb3 <- glm.nb(Canis.lupus ~ Distance_rescaled*rescaled_Time.Diff_new+Site + 
                           sinus_Monitoring.Time.Diff+cosinus_Monitoring.Time.Diff, data = wolf.herb_thin, maxit = 999)

summary(glm.wolf.herb3)

wolf.herb_thin$Site <- relevel(wolf.herb_thin$Site,ref="Yavneel")


interact_plot(model = glm.wolf.herb3, data=wolf.herb_thin, pred = rescaled_Time.Diff_new, modx =Distance_rescaled , partial.residuals = F, jitter = c(0.2), point.size = 3, rug = F, point.shape = F, interval = T,modx.values = "plus-minus")

#no interval:
wolf_inetraction_plot <- interact_plot(model = glm.wolf.herb3, data=wolf.herb_thin, pred = rescaled_Time.Diff_new, modx =Distance_rescaled , partial.residuals = F, jitter = c(0.2), point.size = 3, rug = F, point.shape = F, interval = F,modx.values = "plus-minus")

unique(wolf.herb$rescaled_Time.Diff_new)
```





```{r community analysis - per species models - fox}
fox.herb <- Herb_abundmat[,c(1:17,24:26)]

unique(fox.herb$Site)

names(fox.herb)[names(fox.herb) == 'Vulpes vulpes'] <- 'Vulpes.vulpes'

#the time and distance are significant separately, so model without interaction:

glm.fox.herb <- glm.nb(Vulpes.vulpes ~ Distance_rescaled+rescaled_Time.Diff_new+Site + 
                          sinus_Monitoring.Time.Diff+cosinus_Monitoring.Time.Diff, data = fox.herb, maxit = 999)

summary(glm.fox.herb)

#for distance:
effect_plot_fox.herb <- effect_plot(model = glm.fox.herb, data=fox.herb, 
    pred = Distance_rescaled, interval = T, plot.points = F, jitter = c(0.1,0), 
    point.size = 3, cat.pred.point.size = 4, partial.residuals = F, colors = "Qual1",line.colors = "black") 
effect_plot_fox.herb+ylim(0,40)


#time:
effect_plot_fox.herb2 <- effect_plot(model = glm.fox.herb, data=fox.herb, pred = rescaled_Time.Diff_new, interval = T, plot.points = T, jitter = c(0.1,0), point.size = 3, cat.pred.point.size = 4, partial.residuals = F, colors = "Qual1",line.colors = "black") 
effect_plot_fox.herb+ylim(0,40)

#no dots:
effect_plot_fox.herb2 <- effect_plot(model = glm.fox.herb, data=fox.herb, pred = rescaled_Time.Diff_new, interval = T, plot.points = F, jitter = c(0.1,0), point.size = 3, cat.pred.point.size = 4, partial.residuals = F, colors = "Qual1",line.colors = "black") 

effect_plot_fox.herb2 <- effect_plot(model = glm.fox.herb, data=fox.herb, pred = rescaled_Time.Diff_new, interval = F, plot.points = F, jitter = c(0.1,0), point.size = 3, cat.pred.point.size = 4, partial.residuals = F, colors = "Qual1",line.colors = "black") 

effect_plot(model = glm.fox.herb, data=fox.herb, pred = rescaled_Time.Diff_new, interval = T, plot.points = T, jitter = c(0.1,0), 
            partial.residuals = F, colors = "Qual1") + scale_x_continuous(breaks=c(-1.233466888,-0.601840099,0.028922631,0.66054942,1.291312151), 
                                                                                       labels=c("June 2014", "June 2016", "June 2018", "June 2020", "June 2022"), name = "Time")+
  theme(axis.text=element_text(size=12),axis.title=element_text(size=14,face="bold"))+
  ylab("Red fox abundance")
```


```{r community analysis run model without wolf}

#run the mvabund without wolf (37 indvs and 14 cameras):

Herb_abu_no.wolf <- Herb_abundmat[,c(18,20:24)]
Herb_spp2 <- mvabund(Herb_abu_no.wolf)

mva_herb_nb_no.wolf <- manyglm(Herb_spp2~ Distance_rescaled*rescaled_Time.Diff_new+Site + 
                         sinus_Monitoring.Time.Diff+cosinus_Monitoring.Time.Diff , family = "negative.binomial", data=env_data.herb)

plot.manyglm(mva_herb_nb_no.wolf, which = 1:3)

#model selection:

drop1(mva_herb_nb_no.wolf)

 
#nothing has to be dropped. 

#the final model is mva_herb_nb_no.wolf

anov_mva_herb_nb_no.wolf <- anova(mva_herb_nb_no.wolf, p.uni = "adjusted")
print(anov_mva_herb_nb_no.wolf)

coefp_Herb_no.wolf <- merge(data.table(t(coef(mva_herb_nb_no.wolf)),keep.rownames=TRUE), data.table(t(anov_mva_herb_nb_no.wolf$uni.p), keep.rownames = TRUE), by = "rn") # coefficients are exponentiated because the link function used for poisson is log -> above 1 is positive and below 1 is negative

# add total species abundance
coefp_Herb_no.wolf <- merge(coefp_Herb_no.wolf,as.data.table(colSums(Herb_spp2),keep.rownames = TRUE), by.x = "rn", by.y = "V1")

colnames(coefp_Herb_no.wolf)

colnames(coefp_Herb_no.wolf) <- c("SciName","Intercept.coef","Distance.coef","Time_diff.coef","Site_Karei.Deshe.coef","Site_Natur.coef","Site_Shaal.coef","Site_Yavneel.coef", "sinus_monitoring.time.coef","cosinus_monitoring.time.coef","Distance_Time.coef", "Intercept.p","Distance.p", "Time_diff.p","Site.p", "sin_time.p","Cosin_time.p","Distance_Time.p" ,"species_abundance")

write.csv(coefp_Herb_no.wolf, "../output/batha/coefficients_Herb_no.wolf.csv")

Herb <- all.data.mammals.final.final.for.richness[all.data.mammals.final.final.for.richness$Unit=="Herbaceous and Dwarf Shrub Vegetation",]

unique(Herb$Species)

#to get the numbers of obs and inds:

Herb.data <- all.data.mammals.final.final.without.rare[all.data.mammals.final.final.without.rare$Unit=="Herbaceous and Dwarf Shrub Vegetation",]

unique(Herb.data$Species)
```



```{r community analysis final model without wolf and hyena}
#remove canis lupus and hyena hyena:

library(dplyr)
Herb.data.no.rare <- Herb.data %>% filter(!Species %in% c('Canis lupus','Hyaena hyaena'))
unique(Herb.data.no.rare$Species)

sum(Herb.data.no.rare$Sum)
```


```{r Export Hebrew plots to PDF}
# The plot:

effect_plot(model = glm.fox.herb, data=fox.herb, pred = rescaled_Time.Diff_new, interval = T, plot.points = T, jitter = c(0.1,0), 
            partial.residuals = F, colors = "Qual1",point.size = 5,line.thickness = 2,point.alpha = 0.25,line.colors = "black") + scale_x_continuous(breaks=c(-1.233466888,-0.601840099,0.028922631,0.66054942,1.291312151), 
     labels=c("2014", "2016", "2018", "2020", "2022"))+
  theme(axis.text=element_text(size=18),axis.title=element_text(size=22,face="bold"))+
  ylab("עפש")+xlab("ןמז")



#Red fox plot using function


redfox_shrub <- effect_plot(model = glm.fox.herb, data=fox.herb, pred = rescaled_Time.Diff_new, interval = T, plot.points = T, jitter = c(0.1,0), 
                            partial.residuals = F, colors = "Qual1",point.size = 5,line.thickness = 2,point.alpha = 0.25,line.colors = "black") 
                                                                                                                                                                   
effplot_vulpes_time <- plot_model_effect(P.anal = fox.herb, 
                               m = glm.fox.herb, 
                               eff2plot = "rescaled_Time.Diff_new", 
                               plot_points=FALSE, plot_residuals=FALSE, 
                               export_plot=TRUE, ylabel = NULL, fontname = "Almoni ML v5 AAA",
                               fontsize=22, pdf_width=160, outpath = "../output/batha/")
```

### Session information

```{r}
session_info()
```
