---
title: "Mammal data - Arid Desert"
author: "Talia Gabay & Ron Chen"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#knitr::opts_chunk$set(dev = "cairo_pdf", dev.args=list(cairo_pdf = list(family='Arial')))

#Load packages
library(devtools)
library(rlang)
library(ggplot2)
library(data.table)
library(dplyr)
library(mvabund)
library(kableExtra)
library(lme4)
library(MASS)
library(vegan)
library(jtools)
library(interactions)
library(car)
library(ecoCopula)
library(chron)
library(data.table)
library(readxl)
library(performance)
library(glmmTMB) 
library(emmeans)
library(Cairo)
library(extrafont)

#Load functions
source("functions/plot_alpha_diversity_cycle_num.R")
source("functions/plot_model_effect_mammals.R")
source("functions/plot_interaction_mammals.R")
source_url("https://raw.githubusercontent.com/ronchen1/ron_functions/d2d0bab4dc65f29e53d316ae4098d0469711905b/R/multiplot.R", sha1 = "eef55df1963ba201f57cf44513f566298756fd4f")


```



```{r load data, include=FALSE}
# Abundance table - all data
all.mammal.abundances <- read.csv("Data/mammals.abundance.matrix.TG.csv", header = TRUE, sep =",") %>% as.data.table()

#Incidence table - all data
#all.mammal.incidences <- read.csv("C:/Users/Talia Gabay/OneDrive - #post.bgu.ac.il/Hamaarag_Mammals_analysis_Talia_Copy/Data/mammals.indcidence.matrix.csv", header = #TRUE, sep =",") %>% as.data.table()

#Species names
sci_names <- readxl::read_excel(path = 'Data/Scientific.names.mammals.xlsx') %>% as.data.table()
#Make species list
sci_names_list <- as.character(sort(unique(sci_names$SciName)))
#Change species column names in abundance table 
colnames(all.mammal.abundances)[27:40] <- sci_names_list 
#colnames(all.mammal.abundances)[25:38] <- sci_names_list   
   
#Arid Desert without rare species
#inland.sand.abundance.no.rare <- read.csv("C:/Users/Talia Gabay/OneDrive - #post.bgu.ac.il/Hamaarag_Mammals_analysis_Talia_Copy/Data/mammals.abundance.matrix.inland.no.rare.csv"#, header = TRUE, sep =",") %>% as.data.table()

  P_byplot <- copy(all.mammal.abundances)
  P_byplot[,`:=`(richness = apply(.SD,1,(\(x) sum(x>0))),
                 abundance = apply(.SD,1,sum),
                 gma=apply(.SD,1,(\(x) exp( sum(log(x[x>0])) / sum(x>0) )))),
           .SDcols = sci_names_list]

```

```{r load using function}

P <- P_byplot[grepl("Arid Desert",Unit),][order(Cycle_number,Site)][,Site:=factor(Site)]
```

## Arid Desert - Richness by camera

Monitoring started in 2014 (5 monitoring cycles), Each site as two types of plots - near or far from Agriculture. 
There are 5 sites total - Ein Yahav, Lotan, Paran, Yotvata, Zofar 
Each has 5 sampling points except for Paran (near) which is missing samples from the first year (2014)

Let's first validate the data:

```{r data validation}
#include rare species in analysis
P.anal <- copy(P) # set a fixed variable name for analysis, if want to switch between data WITH rare #species and data WITHOUT rare species then only change once here
#Validate factors and levels
P.anal$Agriculture <- as.factor(P.anal$Agriculture)
print("Agriculture has 2 levels")
print(levels(P.anal$Agriculture))

P.anal$Site <- as.factor(P.anal$Site)
print("Site has 5 levels")
print(levels(P.anal$Site))

P.anal$Transect <- as.factor(P.anal$Transect)
print("Transect has 10 levels")
print(levels(P.anal$Transect))

P.anal$Transect_with_date <- as.factor(P.anal$Transect_with_date)
print("Transect with date has 49 levels - Missing data for Sde Boker near in 2014")
print(levels(P.anal$Transect_with_date))

#distance.rescaled <- scale(all.mammal.abundances$Distance)
#write.csv(distance.rescaled, file = "Distance_agri_rescaled.csv")
```

```{r richness data explore, fig.height = 8, fig.width = 10}
print("RICHNESS WITH RARE SPECIES")
plot_alpha_diversity(P, x_val = "Agriculture", y_val = "richness", ylab_val = "richness", xlab_val = "Agriculture", fill_val = "Agriculture")

col_names <- c("Cycle_number",  "Deployment.id_new",	"Unit",	"Subunit",	"Site",	"Distance_agri_rescaled","Lon",	"Lat", "year", "rescaled_Time.Diff", "sinus_Monitoring.Time.Diff", "cosinus_Monitoring.Time.Diff", "Transect", "Transect_with_date")

abu_by_spp.AridDesert <- all.mammal.abundances[grepl("Arid Desert",Unit),]
spp <- abu_by_spp.AridDesert[,27:40]
#spp <- abu_by_spp.AridDesert[,25:38]
# filter out species with zero counts 
print(colSums(spp))
spp <- spp[,.SD,.SDcols = colSums(spp)>0]
spp <- mvabund(spp)

plot(spp ~ P.anal$Agriculture, overall.main="raw abundances", transformation = "no")

my_cols <- c("blue", "darkgreen")
dotchart(P.anal$richness, ylab = "Agriculture",
         groups = P.anal$Agriculture, gcolor = my_cols,
         color = my_cols[P.anal$Agriculture],
         cex = 0.9,  pch = 1, xlab = "richness")


my_cols <- c("blue", "darkgreen", "red", "orange", "purple")
dotchart(P.anal$richness, ylab = "Site",
         groups = P.anal$Site, gcolor = my_cols,
         color = my_cols[P.anal$Site],
         cex = 0.9,  pch = 1, xlab = "richness")

kable(summary(P.anal[,.(richness, abundance, rescaled_Time.Diff, cosinus_Monitoring.Time.Diff, sinus_Monitoring.Time.Diff, Site, Transect, Transect_with_date, Distance_agri_rescaled)]))
pairs(P.anal[,lapply(X = .SD,FUN = as.numeric),.SDcols=c("sinus_Monitoring.Time.Diff", "cosinus_Monitoring.Time.Diff", "rescaled_Time.Diff","Site","Transect", "Transect_with_date", "Distance_agri_rescaled")])
kable(cor(P.anal[,lapply(X = .SD,FUN = as.numeric),.SDcols=c("sinus_Monitoring.Time.Diff", "cosinus_Monitoring.Time.Diff", "rescaled_Time.Diff","Site","Transect", "Transect_with_date", "Distance_agri_rescaled")]))
meanvar.plot(spp, xlab = "mean abundance of a given species across sites", ylab = "variance of the abundance of a given species across sites")
```

```{r model richness, warning=FALSE}

mdl_r.poiss.int <- glm(data = P.anal, formula = richness ~ Distance_agri_rescaled * rescaled_Time.Diff + cosinus_Monitoring.Time.Diff + sinus_Monitoring.Time.Diff + Site + Transect_with_date, family = poisson)
od <- mdl_r.poiss.int$deviance/mdl_r.poiss.int$df.residual
print("Estimating overdispersion parameter phi: (Res. Dev.)/(n-p) where n=number of observations; p=number of parameters in the model.")
print(paste0('od = ',od))
```

Overdispersion parameter is <1 (underdispersion) and therefore will use Poisson

```{r model selection richness}
m0 <- glm(data = P.anal, formula =  richness ~ Distance_agri_rescaled * rescaled_Time.Diff + cosinus_Monitoring.Time.Diff + sinus_Monitoring.Time.Diff, family = poisson)

me0transectwithdate <- glmer(data = P.anal, formula =   richness ~ Distance_agri_rescaled * rescaled_Time.Diff + cosinus_Monitoring.Time.Diff + sinus_Monitoring.Time.Diff +  (1|Transect_with_date), family = poisson, verbose = 1)

me0transect <- glmer(data = P.anal, formula =   richness ~ Distance_agri_rescaled * rescaled_Time.Diff + cosinus_Monitoring.Time.Diff + sinus_Monitoring.Time.Diff +  (1|Transect), family = poisson, verbose = 1)

me0site <- glmer(data = P.anal, formula =   richness ~ Distance_agri_rescaled * rescaled_Time.Diff + cosinus_Monitoring.Time.Diff + sinus_Monitoring.Time.Diff +  (1|Site), family = poisson, verbose = 1)

```

All GLMERs result in singular fit, will use regular GLM:

```{r model selection richness 2}
m1 <- step(m0)
```

Final model includes distance from agriculture and time difference

```{r model interpretation richness}
summ(m1, exp = TRUE, digits=3)

effect_plot(model = m1, data=P.anal, pred = rescaled_Time.Diff, interval = T, plot.points = F, jitter = c(0.1,0), point.size = 3, cat.pred.point.size = 4, partial.residuals = F, colors = "Qual1", main.title = "Temporal effect (points are partial residuals)")

effect_plot(model = m1, data=P.anal, pred = Distance_agri_rescaled, interval = T, plot.points = F, jitter = c(0.1,0), point.size = 3, cat.pred.point.size = 4, partial.residuals = F, colors = "Qual1", main.title = "Effect of distance from agriculture")

```

Time difference and distance from agriculture significantly affect species richness 

#### community analysis using package MVabund

Table of abundances per species by camera ID

```{r print table with rare species}
abundance_and_cameras <- as.data.table(abu_by_spp.AridDesert[,c(5,27:40)])
dim(abundance_and_cameras)
canisau <- abundance_and_cameras[,2] > 0
sum(canisau)
canislup <- abundance_and_cameras[,3] > 0
sum(canislup)
capra <- abundance_and_cameras[,4] > 0
sum(capra)
Doma <- abundance_and_cameras[,5] > 0
sum(Doma)
Equus <- abundance_and_cameras[,6] > 0
sum(Equus)
Gazellador <- abundance_and_cameras[,7] > 0
sum(Gazellador)
Gazellagaz <- abundance_and_cameras[,8] > 0
sum(Gazellagaz)
Hyaena <- abundance_and_cameras[,9] > 0
sum(Hyaena)
Hystrix <- abundance_and_cameras[,10] > 0
sum(Hystrix)
Lepus <- abundance_and_cameras[,11] > 0
sum(Lepus)
Meles <- abundance_and_cameras[,12] > 0
sum(Meles)
Oryx <- abundance_and_cameras[,13] > 0
sum(Oryx)
Sus <- abundance_and_cameras[,14] > 0
sum(Sus)
Vulpus <- abundance_and_cameras[,15] > 0
sum(Vulpus)
```

303 cameras total. Species with a minimum of 20 cameras - Canis aureus, Canis lupus, Capra nubiana, Gazella dorcas, Hyaena hyaena, Lepus capensis and vulpus vulpus

```{r filtering out rare species}

spp_no_rare <- abu_by_spp.AridDesert[,27:40]
print(colSums(spp_no_rare))
# filter out species with less than 20 individuals (rare species)
spp_no_rare <- spp_no_rare[,.SD,.SDcols = colSums(spp_no_rare)>20]
spp_no_rare <- mvabund(spp_no_rare[,c(-4,-7,-9)])
print(colSums(spp_no_rare))
```

Left with 7 species


```{r community analysis MVabund, fig.height=8, fig.width=10}

env_data <- abu_by_spp.AridDesert[,..col_names]
env_data$Site <- factor(env_data$Site, levels=c("Yotvata","Lotan","Paran", "Zofar", "Ein Yahav"))

mva_m0.po <- manyglm(formula = spp_no_rare ~ Distance_agri_rescaled * rescaled_Time.Diff + cosinus_Monitoring.Time.Diff + sinus_Monitoring.Time.Diff, family = "poisson", data = env_data)

mva_m0.nb <- manyglm(formula = spp_no_rare ~ Distance_agri_rescaled * rescaled_Time.Diff + cosinus_Monitoring.Time.Diff + sinus_Monitoring.Time.Diff, family = "negative.binomial", data = env_data)

aic_dt <- data.table(nb = mva_m0.nb$aic, po=mva_m0.po$aic)
colMeans(aic_dt)
print("POISSON")
plot(mva_m0.po, which=1:3)
print("NEGATIVE BINOMIAL")
plot(mva_m0.nb, which=1:3)
```

prefer negative binomial because standardized residuals are smaller.

```{r community analysis MVabund 2}
# mva_m0 <- mva_m0.po
# mva_m1 <- manyglm(formula = spp_no_rare ~ Distance_agri_rescaled * rescaled_Time.Diff + dunes*rescaled_Time.Diff + Distance_agri_rescaled:dunes + cos_td_rad + sin_td_rad + site, family = "poisson", data = env_data)
# aic_dt$po1 <- mva_m1$aic
mva_m0 <- mva_m0.nb
mva_m1 <- manyglm(formula = spp_no_rare ~ Distance_agri_rescaled * rescaled_Time.Diff + cosinus_Monitoring.Time.Diff + sinus_Monitoring.Time.Diff + Transect_with_date, data = env_data)
aic_dt$nb1 <- mva_m1$aic
colMeans(aic_dt)

mva_m2 <- manyglm(formula = spp_no_rare ~ Distance_agri_rescaled * rescaled_Time.Diff + cosinus_Monitoring.Time.Diff + sinus_Monitoring.Time.Diff + Transect, data = env_data)
aic_dt$nb2 <- mva_m2$aic
colMeans(aic_dt)

mva_m3 <- manyglm(formula = spp_no_rare ~ Distance_agri_rescaled * rescaled_Time.Diff + cosinus_Monitoring.Time.Diff + sinus_Monitoring.Time.Diff + Site, data = env_data)
aic_dt$nb3 <- mva_m2$aic
colMeans(aic_dt)
```

Transect and Site improve the model. 

## Many GLM with Site:

```{r community analysis (site) MVabund model selection}
mva_m0.nb <- manyglm(formula = spp_no_rare ~ Distance_agri_rescaled * rescaled_Time.Diff + cosinus_Monitoring.Time.Diff + sinus_Monitoring.Time.Diff + Site, family = "negative.binomial", data = env_data)
drop1(mva_m0.nb)
```

Final model includes all factors.

```{r community analysis (site) MVabund model selection 2}
# mva_m2 <- manyglm(formula = spp_no_rare ~ Distance_agri_rescaled + dunes*rescaled_Time.Diff + Distance_agri_rescaled:dunes + cos_td_rad + sin_td_rad, family = "poisson", data = env_data)
# mva_m2 <- manyglm(formula = spp_no_rare ~ Distance_agri_rescaled * rescaled_Time.Diff + cosinus_Monitoring.Time.Diff + Site, family = "negative.binomial", data = env_data)
# drop1(mva_m2)
```


```{r community analysis (site) MVabund model selection 3}
# mva_m2 <- manyglm(formula = spp_no_rare ~ Distance_agri_rescaled + dunes + rescaled_Time.Diff + Distance_agri_rescaled:dunes + cos_td_rad + sin_td_rad, family = "poisson", data = env_data)
# mva_m2 <- manyglm(formula = spp_no_rare ~ Distance_agri_rescaled * rescaled_Time.Diff + Site, family = "negative.binomial", data = env_data)
# drop1(mva_m2)
```

```{r community analysis (site) MVabund model validation, fig.height=8, fig.width=10}
plot(mva_m0.nb, which=1:3)
```

```{r community analysis (site) MVabund model interpretation}

#summ_m2 <- summary(mva_m0.nb)
#saveRDS(summ_m2, "output/summ_model_arid_site.RDS")
summ_m2 <- readRDS("output/summ_model_arid_site.RDS")
#anov_m2.uni <- anova(mva_m0.nb, p.uni = "adjusted")
#saveRDS(anov_m2.uni, "output/anova_arid_site.RDS")
anov_m2.uni <- readRDS("output/anova_arid_site.RDS")
print(summ_m2)
print(anov_m2.uni)

```

**Multivariate test shows that all factors (Distance from agriculture, time difference and the interaction between the two, sin and cosin time difference and Site) significantly affect species composition**


```{r community analysis (site) MVabund model interpretation - plot coefficients, error=TRUE}
coefp <- merge(data.table(t(coef(mva_m0.nb)/log(2)),keep.rownames=TRUE), data.table(t(anov_m2.uni$uni.p), keep.rownames = TRUE), by = "rn") # coefficients are exponentiated because the link function used for poisson is log -> above 1 is positive and below 1 is negative
# add total species abundance
coefp <- merge(coefp,as.data.table(colSums(spp_no_rare),keep.rownames = TRUE), by.x = "rn", by.y = "V1")

colnames(coefp) <- c("SciName","Intercept.coef","Distance_agri.coef","Time_diff.coef","cosin_time.coef","sin_time.coef","Site_Lotan.coef","Site_Paran.coef","Site_Zofar.coef","Site_Ein Yahav.coef","Distance_agri_time.coef","Intercept.p","Distance_agri.p","Time_diff.p","cosin_time.p","sin_time.p","Site.p","Distance_agri_time.p","species_abundance")

write.csv(coefp, "coefficients_Arid_Desert.csv")

```

# Per-species models - for Canis aureus only used one Site (Yotvata) and for Capra nubiana only two sites (Yotvata and Lotan) because of many zeros
```{r community analysis (site) Per species models - with interaction}

####Capra nubiana####

only_Capra_data <- P.anal[,c(1:26,29)]
Capra.nubiana.two.sites <- only_Capra_data[Site %in% c("Yotvata", "Lotan")]
glm.Capra.nubiana <- glm.nb(`Capra nubiana` ~ Distance_agri_rescaled * rescaled_Time.Diff + cosinus_Monitoring.Time.Diff + sinus_Monitoring.Time.Diff + Site, data = Capra.nubiana.two.sites, maxit = 999)
coef(glm.Capra.nubiana)
coef(mva_m0.nb)

source("functions/plot_interaction_mammals.R")

plot_model_interaction(P.anal = Capra.nubiana.two.sites, m = glm.Capra.nubiana, eff2plot = "rescaled_Time.Diff", modvar2plot = "Distance_agri_rescaled", plot_points = FALSE, export_plot = TRUE, outpath = "output/arid/", fontname = "Almoni ML v5 AAA")

# capra_interact <- interact_plot(model = glm.Capra.nubiana, data=Capra.nubiana.two.sites, pred = rescaled_Time.Diff, modx = Distance_agri_rescaled, partial.residuals = F, modx.values = "plus-minus", jitter = c(0.2), point.size = 3, rug = F, point.shape = F, interval = T, colors = "Dark2") + scale_x_continuous(breaks=c(-1.233466888,-0.601840099,0.028922631,0.66054942,1.291312151), labels=c("June 2014", "June 2016", "June 2018", "June 2020", "June 2022"), name = "Time")


#Testing whether trends and means are different from zero

#Calculate mean and sd of distance from agriculture
sd_distance_agri <- sd(env_data$Distance_agri_rescaled)
print(sd_distance_agri)
mean_distance_agri <- mean(env_data$Distance_agri_rescaled)
print(mean_distance_agri)

#calculate +/1 1 SD
small_distance <- mean_distance_agri - sd_distance_agri 
large_distance <- mean_distance_agri + sd_distance_agri

#Test whether temporal trend is different from zero
mm_rescaled_time <- emtrends(glm.Capra.nubiana, specs = "Distance_agri_rescaled", var = "rescaled_Time.Diff", type = "response", at = list(Distance_agri_rescaled = c(small_distance,large_distance)))
print(mm_rescaled_time)
test_results_mm_rescaled_time <- test(mm_rescaled_time, null = 0, adjust = "fdr")
print(test_results_mm_rescaled_time)

#Testing whether abundance is higher near or far from agriculture
pairwise_distance <- emmeans(object = glm.Capra.nubiana, ~Distance_agri_rescaled*rescaled_Time.Diff, at = list(Distance_agri_rescaled = c(small_distance,large_distance)))
print(pairwise_distance)
test_results_distance <- test(pairs(pairwise_distance, by="rescaled_Time.Diff"), by=NULL, adjust="fdr")
print(test_results_distance)

####Gazella dorcas####

Gazella.dorcas <- spp_no_rare[,"Gazella.dorcas"]
glm.Gazella.dorcas <- glm.nb(Gazella.dorcas ~ Distance_agri_rescaled * rescaled_Time.Diff + cosinus_Monitoring.Time.Diff + sinus_Monitoring.Time.Diff + Site, data = env_data, maxit = 999)
coef(glm.Gazella.dorcas)
coef(mva_m0.nb)

plot_model_interaction(P.anal = env_data, m = glm.Gazella.dorcas, eff2plot = "rescaled_Time.Diff", modvar2plot = "Distance_agri_rescaled", plot_points = FALSE, export_plot = TRUE, outpath = "output/arid/", fontname = "Almoni ML v5 AAA", legend_position = "bottom")

# interact_plot(model = glm.Gazella.dorcas, data=env_data, pred = rescaled_Time.Diff, modx = Distance_agri_rescaled, partial.residuals = F, modx.values = "plus-minus", jitter = c(0.2), point.size = 3, rug = F, point.shape = F, interval = T, colors = "Dark2") + scale_x_continuous(breaks=c(-1.233466888,-0.601840099,0.028922631,0.66054942,1.291312151), labels=c("June 2014", "June 2016", "June 2018", "June 2020", "June 2022"), name = "Time")

#Test whether temporal trend is different from zero
mm_rescaled_time <- emtrends(glm.Gazella.dorcas, specs = "Distance_agri_rescaled", var = "rescaled_Time.Diff", type = "response", at = list(Distance_agri_rescaled = c(small_distance,large_distance)))
print(mm_rescaled_time)
test_results_mm_rescaled_time <- test(mm_rescaled_time, null = 0, adjust = "fdr")
print(test_results_mm_rescaled_time)

#Testing whether abundance is higher near or far from agriculture
pairwise_distance <- emmeans(object = glm.Gazella.dorcas, ~Distance_agri_rescaled*rescaled_Time.Diff, at = list(Distance_agri_rescaled = c(small_distance,large_distance)))
print(pairwise_distance)
test_results_distance <- test(pairs(pairwise_distance, by="rescaled_Time.Diff"), by=NULL, adjust="fdr")
print(test_results_distance)

```

Capra: 

Temporal trend is significantly different from zero in large distances from agriculture (p < 0.0001) and not significantly different from zero in small distances (p = 0.19)

On average over time, ibex abundance is significantly higher near agriculture (p < 0.0001)



Gazella dorcas:

Temporal trend is significantly different from zero in small distances from agriculture (p < 0.0001) and not significantly different from zero in large distances (p = 0.65)

On average over time, Gazelle abundance is significantly higher near agriculture (p < 0.0001)

```{r community analysis (site) Per species models - no interaction}

####Canis aureus####

only_canis_aur_data <- P.anal[,c(1:27)]
Canis.aureus.only.yotvata <- only_canis_aur_data[Site %in% c("Yotvata")]
Canis.aureus <- spp_no_rare[,"Canis.aureus"]
glm.Canis.aureus <- glm.nb(`Canis aureus` ~ Distance_agri_rescaled + rescaled_Time.Diff + cosinus_Monitoring.Time.Diff + sinus_Monitoring.Time.Diff, data = Canis.aureus.only.yotvata, maxit = 999)
#coef(glm.Canis.aureus)
#coef(mva_m0.nb)

source("functions/plot_model_effect_mammals.R")

plot_model_effect(P.anal = Canis.aureus.only.yotvata, m = glm.Canis.aureus, eff2plot = "rescaled_Time.Diff", plot_points=FALSE, plot_residuals=FALSE, export_plot=TRUE, ylabel = NULL, fontname = "Almoni ML v5 AAA", fontsize=22, pdf_width=160, outpath = "output/arid/")

# effect_plot(model = glm.Canis.aureus, data=Canis.aureus.only.yotvata, pred = rescaled_Time.Diff, interval = T, plot.points = F, jitter = c(0.1,0), point.size = 3, cat.pred.point.size = 4, partial.residuals = F, colors = "Qual1") + ylim(0,5) + scale_x_continuous(breaks=c(-1.233466888,-0.601840099,0.028922631,0.66054942,1.291312151), labels=c("June 2014", "June 2016", "June 2018", "June 2020", "June 2022"), name = "Time")

plot_model_effect(P.anal = Canis.aureus.only.yotvata, m = glm.Canis.aureus, eff2plot = "Distance_agri_rescaled", plot_points=FALSE, plot_residuals=FALSE, export_plot=TRUE, ylabel = NULL, fontname = "Almoni ML v5 AAA", fontsize=22, pdf_width=160, outpath = "output/arid/")

# effect_plot(model = glm.Canis.aureus, data=Canis.aureus.only.yotvata, pred = Distance_agri_rescaled, interval = T, plot.points = F, jitter = c(0.1,0), point.size = 3, cat.pred.point.size = 4, partial.residuals = F, colors = "Qual1") + scale_x_continuous(breaks=c(-0.589201946,0.245068396,1.172035444,2.099002491), labels=c("100", "1000", "2000", "3000"), name = "Distance (m)")
# 

####Canis lupus####

Canis.lupus <- spp_no_rare[,"Canis.lupus"]
glm.Canis.lupus <- glm.nb(Canis.lupus ~ Distance_agri_rescaled + rescaled_Time.Diff + cosinus_Monitoring.Time.Diff + sinus_Monitoring.Time.Diff + Site, data = env_data, maxit = 999)
coef(glm.Canis.lupus)
coef(mva_m0.nb)

plot(glm.Canis.lupus)

plot_model_effect(P.anal = env_data, m = glm.Canis.lupus, eff2plot = "rescaled_Time.Diff", plot_points=FALSE, plot_residuals=FALSE, export_plot=TRUE, ylabel = NULL, fontname = "Almoni ML v5 AAA", fontsize=22, pdf_width=160, outpath = "output/arid/")

# effect_plot(model = glm.Canis.lupus, data=env_data, pred = rescaled_Time.Diff, interval = T, plot.points = F, jitter = c(0.1,0), point.size = 3, cat.pred.point.size = 4, partial.residuals = F, colors = "Qual1") + scale_x_continuous(breaks=c(-1.233466888,-0.601840099,0.028922631,0.66054942,1.291312151), labels=c("June 2014", "June 2016", "June 2018", "June 2020", "June 2022"), name = "Time")

```

n = 303 (total number of cameras)

total abundance = 3035 (10 species)

total abundance without rare species = 2933 (7 species)

### Session information

```{r}
session_info()
```

