---
title: "Mammal data - Mediterranean Maquis - no interaction with subunit"
author: "Ron Chen & Talia Gabay"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)
#knitr::opts_chunk$set(dev = "cairo_pdf", dev.args=list(cairo_pdf = list(family='Arial')))

#Load packages
library(devtools)
library(rlang)
library(ggplot2)
library(data.table)
library(dplyr)
library(mvabund)
library(kableExtra)
library(lme4)
library(MASS)
library(vegan)
library(jtools)
library(interactions)
library(car)
library(ecoCopula)
library(chron)
library(data.table)
library(readxl)
library(performance)
library(Cairo)
library(emmeans)
library(extrafont)

#Load functions
source("functions/plot_alpha_diversity_cycle_num.R")
source("functions/plot_model_effect_mammals_Maquis_and_Forest.R")
source("functions/plot_interaction_mammals_Maquis_and_Forest.R")
source_url("https://raw.githubusercontent.com/ronchen1/ron_functions/d2d0bab4dc65f29e53d316ae4098d0469711905b/R/multiplot.R", sha1 = "eef55df1963ba201f57cf44513f566298756fd4f")

```



```{r load data, include=FALSE}
# Abundance table - all data
all.mammal.abundances <- read.csv("data/mammals.abundance.matrix.TG.csv", header = TRUE, sep =",") %>% as.data.table()


#Species names
sci_names <- readxl::read_excel(path = 'data/Scientific.names.mammals.xlsx') %>% as.data.table()
#Make species list
sci_names_list <- as.character(sort(unique(sci_names$SciName)))
#Change species column names in abundance table 
colnames(all.mammal.abundances)[27:40] <- sci_names_list 
#colnames(all.mammal.abundances)[25:38] <- sci_names_list   
   
all.mammal.abundances <- all.mammal.abundances[,-30]
 
  P_byplot <- copy(all.mammal.abundances)
  P_byplot[,`:=`(richness = apply(.SD,1,(\(x) sum(x>0))),
                 abundance = apply(.SD,1,sum),
                 gma=apply(.SD,1,(\(x) exp( sum(log(x[x>0])) / sum(x>0) )))),
           .SDcols = sci_names_list[-4]]

```

```{r load using function}

P <- P_byplot[grepl("Mediterranean Maquis",Unit),][order(Cycle_number,Site)][,Site:=factor(Site)]
```

## Mediterranean Maquis - Richness by camera

### Monitoring started in 2014 (5 monitoring cycles), Each site has two types of plots - far or near settlements
### There are 15 sites total, each site has 5 sampling points 

Let's first validate the data:

```{r data validation}
#include rare species in analysis
P.anal <- copy(P) # set a fixed variable name for analysis, if want to switch between data WITH rare #species and data WITHOUT rare species then only change once here
#Validate factors and levels
P.anal$Settlements <- as.factor(P.anal$Settlements)
print("Settlements has 2 levels")
print(levels(P.anal$Settlements))
P.anal$Subunit <- as.factor(P.anal$Subunit)
print("Subunit has 3 levels")
print(levels(P.anal$Subunit))
P.anal$Site <- as.factor(P.anal$Site)
print("Site has 15 levels")
print(levels(P.anal$Site))
P.anal$Transect <- as.factor(P.anal$Transect)
print("Transect has 30 levels")
print(levels(P.anal$Transect))
P.anal$Transect_with_date <- as.factor(P.anal$Transect_with_date)
print("Transect with date has 150 levels")
print(levels(P.anal$Transect_with_date))

```
```{r richness data explore, fig.height = 8, fig.width = 10}
print("RICHNESS WITH RARE SPECIES")
plot_alpha_diversity(P, x_val = "Subunit", y_val = "richness", ylab_val = "richness", xlab_val = "Subunit", fill_val = "Subunit")

col_names <- c("Cycle_number",  "Deployment.id_new",	"Unit",	"Subunit",	"Site", "Settlements",	"Distance_rescaled", "Lon",	"Lat", "year", "rescaled_Time.Diff", "sinus_Monitoring.Time.Diff", "cosinus_Monitoring.Time.Diff", "Transect", "Transect_with_date")

abu_by_spp.Maquis <- all.mammal.abundances[grepl("Mediterranean Maquis",Unit),]
spp <- abu_by_spp.Maquis[,c(27:39)]

# filter out species with zero counts 
print(colSums(spp))
spp <- spp[,.SD,.SDcols = colSums(spp)>0]
spp <- mvabund(spp)

plot(spp ~ P.anal$Subunit, overall.main="raw abundances", transformation = "no")

my_cols <- c("blue", "darkgreen", "red")
dotchart(P.anal$richness, ylab = "Subunit",
         groups = P.anal$Subunit, gcolor = my_cols,
         color = my_cols[P.anal$Subunit],
         cex = 0.9,  pch = 1, xlab = "richness")

kable(summary(P.anal[,.(richness, abundance, Subunit, rescaled_Time.Diff, cosinus_Monitoring.Time.Diff, sinus_Monitoring.Time.Diff, Site, Transect, Transect_with_date, Distance_rescaled)]))
pairs(P.anal[,lapply(X = .SD,FUN = as.numeric),.SDcols=c("Subunit","sinus_Monitoring.Time.Diff", "cosinus_Monitoring.Time.Diff", "rescaled_Time.Diff","Site","Transect", "Transect_with_date", "Distance_rescaled")])
kable(cor(P.anal[,lapply(X = .SD,FUN = as.numeric),.SDcols=c("Subunit","sinus_Monitoring.Time.Diff", "cosinus_Monitoring.Time.Diff", "rescaled_Time.Diff","Site","Transect", "Transect_with_date", "Distance_rescaled")]))
meanvar.plot(spp, xlab = "mean abundance of a given species across sites", ylab = "variance of the abundance of a given species across sites")

```

### Community analysis using package MVabund


Table of abundances per species by camera ID

```{r print table with rare species}
abundance_and_cameras <- as.data.table(abu_by_spp.Maquis[,c(5,27:39)])
dim(abundance_and_cameras)
canisau <- abundance_and_cameras[,2] > 0
sum(canisau)
canislup <- abundance_and_cameras[,3] > 0
sum(canislup)
capra <- abundance_and_cameras[,4] > 0
sum(capra)
Equus <- abundance_and_cameras[,5] > 0
sum(Equus)
Gazellador <- abundance_and_cameras[,6] > 0
sum(Gazellador)
Gazellagaz <- abundance_and_cameras[,7] > 0
sum(Gazellagaz)
Hyaena <- abundance_and_cameras[,8] > 0
sum(Hyaena)
Hystrix <- abundance_and_cameras[,9] > 0
sum(Hystrix)
Lepus <- abundance_and_cameras[,10] > 0
sum(Lepus)
Meles <- abundance_and_cameras[,11] > 0
sum(Meles)
Oryx <- abundance_and_cameras[,12] > 0
sum(Oryx)
Sus <- abundance_and_cameras[,13] > 0
sum(Sus)
Vulpes <- abundance_and_cameras[,14] > 0
sum(Vulpes)
```

1117 cameras. Species with a minimum of 20 cameras - Canis aureus, Canis lupus, Gazella gazella, Hyaena hyaena, Hystrix indica, Meles meles, Sus scorfa and Vulpes vulpes

```{r filtering out rare species}

spp_no_rare <- abu_by_spp.Maquis[,27:39]
print(colSums(spp_no_rare))
# filter out species with less than 20 individuals (rare species)
spp_no_rare <- spp_no_rare[,.SD,.SDcols = colSums(spp_no_rare)>=20]
#spp_no_rare <- spp_no_rare[,c(-2,-6)]
spp_no_rare <- mvabund(spp_no_rare)
print(colSums(spp_no_rare))

#Sus scorfa abundance across subunit, sites and projectid
my_cols <- c("blue", "darkgreen", "red")
dotchart(P.anal$`Sus scrofa`, ylab = "Subunit",
         groups = P.anal$Subunit, gcolor = my_cols,
         color = my_cols[P.anal$Subunit],
         cex = 0.9,  pch = 1, xlab = "Abundance")

P.anal$Project.ID <- as.factor(P.anal$Project.ID)
my_cols <- c("blue", "darkgreen", "red", "purple", "darkorange")
dotchart(P.anal$`Sus scrofa`, ylab = "Project.ID",
         groups = P.anal$Project.ID, gcolor = my_cols,
         color = my_cols[P.anal$Project.ID],
         cex = 0.9,  pch = 1, xlab = "Abundance")

fun_color_range <- colorRampPalette(c("#1b98e0", "red"))
my_cols <- fun_color_range(15)    
dotchart(P.anal$`Sus scrofa`, ylab = "Site",
         groups = P.anal$Site, gcolor = my_cols,
         color = my_cols[P.anal$Site],
         cex = 0.9,  pch = 1, xlab = "Abundance")

```

Species with 20 individuals or more - Canis aureus, Canis lupus, Gazella gazella, Hyaena hyaena, Hystrix indica, Meles meles, Sus scorfa and Vulpes vulpes

Left with 8 species

```{r community analysis MVabund, fig.height=8, fig.width=10}

env_data <- abu_by_spp.Maquis[,..col_names]

mva_m0.po <- manyglm(formula = spp_no_rare ~ Distance_rescaled * rescaled_Time.Diff + cosinus_Monitoring.Time.Diff + sinus_Monitoring.Time.Diff, family = "poisson", data = env_data)

mva_m0.nb <- manyglm(formula = spp_no_rare ~ Distance_rescaled * rescaled_Time.Diff + cosinus_Monitoring.Time.Diff + sinus_Monitoring.Time.Diff, family = "negative.binomial", data = env_data)

aic_dt <- data.table(nb = mva_m0.nb$aic, po=mva_m0.po$aic)
colMeans(aic_dt)
print("POISSON")
plot(mva_m0.po, which=1:3)
print("NEGATIVE BINOMIAL")
plot(mva_m0.nb, which=1:3)
```

prefer negative binomial because standardized residuals are smaller.

```{r community analysis MVabund 2}

# mva_m0 <- mva_m0.po
# mva_m1 <- manyglm(formula = spp_no_rare ~ Distance_rescaled * rescaled_Time.Diff + dunes*rescaled_Time.Diff + Distance_rescaled:dunes + cos_td_rad + sin_td_rad + site, family = "poisson", data = env_data)
# aic_dt$po1 <- mva_m1$aic

mva_m0 <- mva_m0.nb
mva_m1 <- manyglm(formula = spp_no_rare ~ Distance_rescaled * rescaled_Time.Diff + cosinus_Monitoring.Time.Diff + sinus_Monitoring.Time.Diff + Site + Subunit, data = env_data)
aic_dt$nb1 <- mva_m1$aic
colMeans(aic_dt)

mva_m2 <- manyglm(formula = spp_no_rare ~ Distance_rescaled * rescaled_Time.Diff + cosinus_Monitoring.Time.Diff + sinus_Monitoring.Time.Diff + Transect, data = env_data)
aic_dt$nb2 <- mva_m2$aic
colMeans(aic_dt)

mva_m3 <- manyglm(formula = spp_no_rare ~ Distance_rescaled * rescaled_Time.Diff + cosinus_Monitoring.Time.Diff + sinus_Monitoring.Time.Diff + Site, data = env_data)
aic_dt$nb3 <- mva_m2$aic
colMeans(aic_dt)

mva_m4 <- manyglm(formula = spp_no_rare ~ Distance_rescaled * rescaled_Time.Diff + cosinus_Monitoring.Time.Diff + sinus_Monitoring.Time.Diff + Subunit, data = env_data)
aic_dt$nb4 <- mva_m2$aic
colMeans(aic_dt)
```

Will proceed with subunit:


# Many GLM with subunit

```{r community analysis (subunit) MVabund model selection}
mva_m0.nb <- manyglm(formula = spp_no_rare ~ Distance_rescaled * rescaled_Time.Diff + cosinus_Monitoring.Time.Diff + sinus_Monitoring.Time.Diff + Subunit, family = "negative.binomial", data = env_data)
drop1(mva_m0.nb)

```

Dropped sin time

```{r community analysis (subunit) MVabund model selection 2}
# mva_m2 <- manyglm(formula = spp_no_rare ~ Distance_rescaled + dunes*rescaled_Time.Diff + Distance_rescaled:dunes + cos_td_rad + sin_td_rad, family = "poisson", data = env_data)
mva_m2 <- manyglm(formula = spp_no_rare ~ Distance_rescaled * rescaled_Time.Diff + cosinus_Monitoring.Time.Diff + Subunit, family = "negative.binomial", data = env_data)
drop1(mva_m2)
```

 
```{r community analysis (subunit) MVabund model selection 3}
# mva_m2 <- manyglm(formula = spp_no_rare ~ Distance_rescaled + dunes + rescaled_Time.Diff + Distance_rescaled:dunes + cos_td_rad + sin_td_rad, family = "poisson", data = env_data)
# mva_m2 <- manyglm(formula = spp_no_rare ~ Distance_rescaled + cosinus_Monitoring.Time.Diff + sinus_Monitoring.Time.Diff + Site, family = "negative.binomial", data = env_data)
# drop1(mva_m2)
```


```{r community analysis (subunit) MVabund model selection 4}
# mva_m2 <- manyglm(formula = spp_no_rare ~ Distance_rescaled + dunes + rescaled_Time.Diff + Distance_rescaled:dunes + cos_td_rad + sin_td_rad, family = "poisson", data = env_data)
# mva_m2 <- manyglm(formula = spp_no_rare ~ Distance_rescaled + sinus_Monitoring.Time.Diff + Site, family = "negative.binomial", data = env_data)
# drop1(mva_m2)
```

## Model includes Subunit, cosin time difference and the interaction between time and distance from settlement (but see below, final model without interaction)

```{r community analysis (subunit) MVabund model selection 5}
# mva_m2 <- manyglm(formula = spp_no_rare ~ Dunes*rescaled_Time.Diff + sinus_Monitoring.Time.Diff + Transect_with_date, data = env_data)
 # drop1(mva_m2)
```

```{r community analysis (subunit) MVabund model validation, fig.height=8, fig.width=10}
plot(mva_m2, which=1:3)
```

```{r community analysis (subunit) MVabund model interpretation}

#summ_m2 <- summary(mva_m2)
#saveRDS(summ_m2, "output/summ_maquis_model_no_subunit_interaction.RDS")
summ_m2 <- readRDS("output/summ_maquis_model_no_subunit_interaction.RDS")
print(summ_m2)
#anov_m2.uni <- anova(mva_m2, p.uni = "adjusted")
#saveRDS(anov_m2.uni, "output/anova_maquis_model_no_subunit_interaction.RDS")
anov_m2.uni <- readRDS("output/anova_maquis_model_no_subunit_interaction.RDS")
print(anov_m2.uni)

```

## Multivariate test shows that subunit, time, distance from settlements and timeXdistance affect species composition
### However, the interaction of time by distance from settlements is not significant for any of the species (H. indica is nearly significant). Therefore the manyglm model will be re-run without this interaction.

```{r community analysis without interaction, , fig.height=8, fig.width=10}
mva_m3 <- manyglm(formula = spp_no_rare ~ Distance_rescaled + rescaled_Time.Diff + cosinus_Monitoring.Time.Diff + Subunit, family = "negative.binomial", data = env_data)
plot(mva_m3, which=1:3)

# summ_m3 <- summary(mva_m3)
# saveRDS(summ_m3, "output/summ_maquis_model_no_subunit_interaction_no_distXtime_interaction.RDS")
summ_m3 <- readRDS("output/summ_maquis_model_no_subunit_interaction_no_distXtime_interaction.RDS")
print(summ_m3)

# anov_m3.uni <- anova(mva_m3, p.uni = "adjusted")
# saveRDS(anov_m3.uni, "output/anova_maquis_model_no_subunit_interaction_no_distXtime_interaction.RDS")
anov_m3.uni <- readRDS("output/anova_maquis_model_no_subunit_interaction_no_distXtime_interaction.RDS")
print(anov_m3.uni)

```

```{r community analysis (subunit) MVabund model interpretation - plot coefficients, error=TRUE}
coefp <- merge(data.table(t(coef(mva_m3)/log(2)),keep.rownames=TRUE), data.table(t(anov_m3.uni$uni.p), keep.rownames = TRUE), by = "rn") # coefficients are exponentiated because the link function used for poisson is log -> above 1 is positive and below 1 is negative
# add total species abundance
coefp <- merge(coefp,as.data.table(colSums(spp_no_rare),keep.rownames = TRUE), by.x = "rn", by.y = "V1")

colnames(coefp) <- c("SciName","Intercept.coef","Distance.coef","time_diff.coef","Cosin_time.coef","Galilee.coef","Judea.coef","Intercept.p","Distance.p","time_diff.p","Cosin_time.p","Subunit.p","Species_abundance")

write.csv(coefp, "coefficients_Maquis_no_subunit_interaction.csv")

```



### species with statistically significant coefficients are in red:
 - species affected by distance from settlements are C. aureus, G. gazella, H. hyaena and V. vulpes.
 - Only H. indica has temporal trend.
 - all but V. vulpes are affected by subunit.
 - none have significant interaction between distance from settlements and time.

```{r}
cs_dt <- coefp[,c(1,3:7)]
cs_dt$Distance.coef <- cell_spec(cs_dt$Distance.coef, color = ifelse(coefp$Distance.p[1:8]<0.05,"red","black"))
cs_dt$time_diff.coef <- cell_spec(cs_dt$time_diff.coef, color = ifelse(coefp$time_diff.p[1:8]<0.05,"red","black"))
cs_dt$Cosin_time.coef <- cell_spec(cs_dt$Cosin_time.coef, color = ifelse(coefp$Cosin_time.p[1:8]<0.05,"red","black"))
cs_dt$Galilee.coef <- cell_spec(cs_dt$Galilee.coef, color = ifelse(coefp$Subunit.p[1:8]<0.05,"red","black"))
cs_dt$Judea.coef <- cell_spec(cs_dt$Judea.coef, color = ifelse(coefp$Subunit.p[1:8]<0.05,"red","black"))
# cs_dt$Distance_time.coef <- cell_spec(cs_dt$Distance_time.coef, color = ifelse(coefp$Distance_time.p[1:8]<0.05,"red","black"))
kbl(cs_dt, escape = F)
```


```{r community analysis (subunit) - per species models - no interaction term for distance X time}

####Canis aureus####

Canis.aureus <- spp_no_rare[,"Canis.aureus"]
glm.Canis.aureus <- glm.nb(formula = Canis.aureus ~ Distance_rescaled + rescaled_Time.Diff + cosinus_Monitoring.Time.Diff + Subunit, data = env_data)
print(summary(glm.Canis.aureus))
coef(glm.Canis.aureus)
coef(mva_m3)

plot_model_effect(P.anal = env_data, m = glm.Canis.aureus, eff2plot = "Distance_rescaled", plot_points=FALSE, plot_residuals=FALSE, export_plot=TRUE, ylabel = NULL, fontname = "Almoni ML v5 AAA", fontsize=22, pdf_width=160, outpath = "output/maquis/")


####Hystrix indica####

Hystrix.indica <- spp_no_rare[,"Hystrix.indica"]
glm.Hystrix.indica <- glm.nb(formula = Hystrix.indica ~ Distance_rescaled + rescaled_Time.Diff + cosinus_Monitoring.Time.Diff + Subunit, data = env_data)
coef(glm.Hystrix.indica)
coef(mva_m3)

plot_model_effect(P.anal = env_data, m = glm.Hystrix.indica, eff2plot = "rescaled_Time.Diff", plot_points=FALSE, plot_residuals=FALSE, export_plot=TRUE, ylabel = NULL, fontname = "Almoni ML v5 AAA", fontsize=22, pdf_width=160, outpath = "output/maquis/")


####Gazella gazella####

Gazella.gazella <- spp_no_rare[,"Gazella.gazella"]
glm.Gazella.gazella <- glm.nb(formula = Gazella.gazella ~ Distance_rescaled + rescaled_Time.Diff + cosinus_Monitoring.Time.Diff + Subunit, data = env_data)
print(summary(glm.Gazella.gazella))
coef(glm.Gazella.gazella)
coef(mva_m3)

plot_model_effect(P.anal = env_data, m = glm.Gazella.gazella, eff2plot = "rescaled_Time.Diff", plot_points=FALSE, plot_residuals=FALSE, export_plot=TRUE, ylabel = NULL, fontname = "Almoni ML v5 AAA", fontsize=22, pdf_width=160, outpath = "output/maquis/")
plot_model_effect(P.anal = env_data, m = glm.Gazella.gazella, eff2plot = "Distance_rescaled", plot_points=FALSE, plot_residuals=FALSE, export_plot=TRUE, ylabel = NULL, fontname = "Almoni ML v5 AAA", fontsize=22, pdf_width=160, outpath = "output/maquis/")

####Hyaena hyaena####

Hyaena.hyaena <- spp_no_rare[,"Hyaena.hyaena"]
glm.Hyaena.hyaena <- glm.nb(formula = Hyaena.hyaena ~ Distance_rescaled + rescaled_Time.Diff + cosinus_Monitoring.Time.Diff + Subunit, data = env_data)
print(summary(glm.Hyaena.hyaena))
coef(glm.Hyaena.hyaena)
coef(mva_m3)

plot_model_effect(P.anal = env_data, m = glm.Hyaena.hyaena, eff2plot = "Distance_rescaled", plot_points=FALSE, plot_residuals=FALSE, export_plot=TRUE, ylabel = NULL, fontname = "Almoni ML v5 AAA", fontsize=22, pdf_width=160, outpath = "output/maquis/", y_cutoff = 2.5)

####Vulpes vulpes####

Vulpes.vulpes <- spp_no_rare[,"Vulpes.vulpes"]
glm.Vulpes.vulpes <- glm.nb(formula = Vulpes.vulpes ~ Distance_rescaled + rescaled_Time.Diff + cosinus_Monitoring.Time.Diff + Subunit, data = env_data)
print(summary(glm.Vulpes.vulpes))
coef(glm.Vulpes.vulpes)
coef(mva_m3)

plot_model_effect(P.anal = env_data, m = glm.Vulpes.vulpes, eff2plot = "Distance_rescaled", plot_points=FALSE, plot_residuals=FALSE, export_plot=TRUE, ylabel = NULL, fontname = "Almoni ML v5 AAA", fontsize=22, pdf_width=160, outpath = "output/maquis/")

```

### Analyze two species with interaction between time and distance from settlements

**Hystrix indica**

```{r community analysis (subunit) - per species models (with interaction between time and distance from settlements - H. indica)}
#####Hystrix indica####

Hystrix.indica <- spp_no_rare[,"Hystrix.indica"]
glm.Hystrix.indica <- glm.nb(formula = Hystrix.indica ~ Distance_rescaled * rescaled_Time.Diff + cosinus_Monitoring.Time.Diff + Subunit, data = env_data)
coef(glm.Hystrix.indica)
coef(mva_m3)

plot_model_interaction(P.anal = env_data, m = glm.Hystrix.indica, eff2plot = "rescaled_Time.Diff", modvar2plot = "Distance_rescaled", plot_points=FALSE, plot_residuals=FALSE, export_plot=TRUE, ylabel = NULL, fontname = "Almoni ML v5 AAA",
fontsize=22, pdf_width=160, outpath = "output/maquis/")


sd_distance <- sd(env_data$Distance_rescaled)
print(sd_distance)
mean_distance <- mean(env_data$Distance_rescaled)
print(mean_distance)

#calculate +/1 1 SD
small_distance <- mean_distance - sd_distance 
large_distance <- mean_distance + sd_distance

#Test whether temporal trend is different from zero
mm_rescaled_time <- emtrends(glm.Hystrix.indica, specs = "Distance_rescaled", var = "rescaled_Time.Diff", type = "response", at = list(Distance_rescaled = c(small_distance,large_distance)))
print(mm_rescaled_time)
test_results_mm_rescaled_time <- test(mm_rescaled_time, null = 0, adjust = "fdr")
print(test_results_mm_rescaled_time)

#Testing whether abundance is higher near or far from settlements
pairwise_distance <- emmeans(object = glm.Hystrix.indica, ~ Distance_rescaled*rescaled_Time.Diff, at = list(Distance_rescaled = c(small_distance,large_distance)))
print(pairwise_distance)
test_results_distance <- test(pairs(pairwise_distance, by="rescaled_Time.Diff"), by=NULL, adjust="fdr")
print(test_results_distance)
```

**Hyaena hyaena**

```{r community analysis (subunit) - per species models (with interaction between time and distance from settlements) - H.hyaena}
#####Hyaena hyaena####

Hyaena.hyaena <- spp_no_rare[,"Hyaena.hyaena"]
glm.Hyaena.hyaena <- glm.nb(formula = Hyaena.hyaena ~ Distance_rescaled * rescaled_Time.Diff + cosinus_Monitoring.Time.Diff + Subunit, data = env_data)
coef(glm.Hyaena.hyaena)
coef(mva_m3)

plot_model_interaction(P.anal = env_data, m = glm.Hyaena.hyaena, eff2plot = "rescaled_Time.Diff", modvar2plot = "Distance_rescaled", plot_points=FALSE, plot_residuals=FALSE, export_plot=TRUE, ylabel = NULL, fontname = "Almoni ML v5 AAA",
fontsize=22, pdf_width=160, outpath = "output/maquis/")


sd_distance <- sd(env_data$Distance_rescaled)
print(sd_distance)
mean_distance <- mean(env_data$Distance_rescaled)
print(mean_distance)

#calculate +/1 1 SD
small_distance <- mean_distance - sd_distance 
large_distance <- mean_distance + sd_distance

#Test whether temporal trend is different from zero
mm_rescaled_time <- emtrends(glm.Hyaena.hyaena, specs = "Distance_rescaled", var = "rescaled_Time.Diff", type = "response", at = list(Distance_rescaled = c(small_distance,large_distance)))
print(mm_rescaled_time)
test_results_mm_rescaled_time <- test(mm_rescaled_time, null = 0, adjust = "fdr")
print(test_results_mm_rescaled_time)

#Testing whether abundance is higher near or far from settlements
pairwise_distance <- emmeans(object = glm.Hyaena.hyaena, ~ Distance_rescaled*rescaled_Time.Diff, at = list(Distance_rescaled = c(small_distance,large_distance)))
print(pairwise_distance)
test_results_distance <- test(pairs(pairwise_distance, by="rescaled_Time.Diff"), by=NULL, adjust="fdr")
print(test_results_distance)

```

### Session information

```{r}
session_info()
```

