---
title: "Mammal data - Planted Conifer Forests"
author: "Talia Gabay & Ron Chen"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE)
#knitr::opts_chunk$set(dev = "cairo_pdf", dev.args=list(cairo_pdf = list(family='Arial')))



#Load packages
library(devtools)
library(rlang)
library(ggplot2)
library(data.table)
library(dplyr)
library(mvabund)
library(kableExtra)
library(lme4)
library(MASS)
library(vegan)
library(jtools)
library(interactions)
library(car)
library(ecoCopula)
library(chron)
library(data.table)
library(readxl)
library(performance)
library(emmeans)
library(extrafont)

#Load functions
source("functions/plot_alpha_diversity_cycle_num.R")
source("functions/plot_model_effect_mammals_Maquis_and_Forest.R")
source("functions/plot_interaction_mammals_Maquis_and_Forest.R")
source_url("https://raw.githubusercontent.com/ronchen1/ron_functions/d2d0bab4dc65f29e53d316ae4098d0469711905b/R/multiplot.R", sha1 = "eef55df1963ba201f57cf44513f566298756fd4f")

```



```{r load data, include=FALSE}
# Abundance table - all data
all.mammal.abundances <- read.csv("Data/mammals.abundance.matrix.TG.csv", header = TRUE, sep =",") %>% as.data.table()


#Species names
sci_names <- readxl::read_excel(path = 'Data/Scientific.names.mammals.xlsx') %>% as.data.table()
#Make species list
sci_names_list <- as.character(sort(unique(sci_names$SciName)))
#Change species column names in abundance table 
colnames(all.mammal.abundances)[27:40] <- sci_names_list 
#colnames(all.mammal.abundances)[25:38] <- sci_names_list   
   
all.mammal.abundances <- all.mammal.abundances[,-30]
 
  P_byplot <- copy(all.mammal.abundances)
  P_byplot[,`:=`(richness = apply(.SD,1,(\(x) sum(x>0))),
                 abundance = apply(.SD,1,sum),
                 gma=apply(.SD,1,(\(x) exp( sum(log(x[x>0])) / sum(x>0) )))),
           .SDcols = sci_names_list[-4]]

```

```{r load using function}

P <- P_byplot[grepl("Planted Conifer Forests",Unit),][order(Cycle_number,Site)][,Site:=factor(Site)]
```

## Planted Conifer Forests - Richness by camera

### Monitoring started in 2014 (5 monitoring cycles), Each site has only one plot that is far from Settlements. 
### There are 15 sites total, each site has 5 sampling points 

Let's first validate the data:

```{r data validation}
#include rare species in analysis
P.anal <- copy(P) # set a fixed variable name for analysis, if want to switch between data WITH rare #species and data WITHOUT rare species then only change once here
#Validate factors and levels
P.anal$Settlements <- as.factor(P.anal$Settlements)
print("Settlements has 1 level")
print(levels(P.anal$Settlements))
P.anal$Subunit <- as.factor(P.anal$Subunit)
print("Subunit has 3 levels")
print(levels(P.anal$Subunit))
P.anal$Site <- as.factor(P.anal$Site)
print("Site has 15 levels")
print(levels(P.anal$Site))
P.anal$Transect <- as.factor(P.anal$Transect)
print("Transect has 15 levels")
print(levels(P.anal$Transect))
P.anal$Transect_with_date <- as.factor(P.anal$Transect_with_date)
print("Transect with date has 75 levels")
print(levels(P.anal$Transect_with_date))

```

```{r richness data explore, fig.height = 8, fig.width = 10}
print("RICHNESS WITH RARE SPECIES")
plot_alpha_diversity(P, x_val = "Subunit", y_val = "richness", ylab_val = "richness", xlab_val = "Subunit", fill_val = "Subunit")

col_names <- c("Cycle_number",  "Deployment.id_new",	"Unit",	"Subunit",	"Site", "Settlements",	"Distance_rescaled", "Lon",	"Lat", "year", "rescaled_Time.Diff", "sinus_Monitoring.Time.Diff", "cosinus_Monitoring.Time.Diff", "Transect", "Transect_with_date")

my_cols <- c("blue", "darkgreen", "red")
dotchart(P.anal$richness, ylab = "Subunit",
         groups = P.anal$Subunit, gcolor = my_cols,
         color = my_cols[P.anal$Subunit],
         cex = 0.9,  pch = 1, xlab = "richness")
```

Explore abundances:

```{r}
abu_by_spp.ConiferForests <- all.mammal.abundances[grepl("Planted Conifer Forests",Unit),]
spp <- abu_by_spp.ConiferForests[,c(27:39)]

# filter out species with zero counts 
print(colSums(spp))
spp <- spp[,.SD,.SDcols = colSums(spp)>0]
spp <- mvabund(spp)

try(plot(spp ~ P.anal$Subunit, overall.main="raw abundances", transformation = "no"))
```

Sus scrofa only:

```{r abundance data explore, fig.height = 8, fig.width = 10}
#Sus scorfa abundance across subunit, sites and projectid
my_cols <- c("blue", "darkgreen", "red")
dotchart(P.anal$`Sus scrofa`, ylab = "Subunit",
         groups = P.anal$Subunit, gcolor = my_cols,
         color = my_cols[P.anal$Subunit],
         cex = 0.9,  pch = 1, xlab = "Abundance")

P.anal$Project.ID <- as.factor(P.anal$Project.ID)
my_cols <- c("blue", "darkgreen", "red", "purple", "darkorange")
dotchart(P.anal$`Sus scrofa`, ylab = "Project.ID",
         groups = P.anal$Project.ID, gcolor = my_cols,
         color = my_cols[P.anal$Project.ID],
         cex = 0.9,  pch = 1, xlab = "Abundance")

fun_color_range <- colorRampPalette(c("#1b98e0", "red"))
my_cols <- fun_color_range(15)    
dotchart(P.anal$`Sus scrofa`, ylab = "Site",
         groups = P.anal$Site, gcolor = my_cols,
         color = my_cols[P.anal$Site],
         cex = 0.9,  pch = 1, xlab = "Abundance")
```

Fix outlier observation of 186 wild boar, by setting the count to be equal to the highest count observed.

```{r fix sus scrofa outlier}
maxI <- which(spp[,"Sus.scrofa"]>100)
spp[maxI,"Sus.scrofa"] <- max(spp[-maxI,"Sus.scrofa"])
P.anal[maxI,`Sus scrofa`:=max(spp[-maxI,"Sus.scrofa"])]
abu_by_spp.ConiferForests[maxI,`Sus scrofa`:=max(spp[-maxI,"Sus.scrofa"])]
try(plot(spp ~ P.anal$Subunit, overall.main="raw abundances", transformation = "no"))
```

Fix outlier observations of canis lupus, gazella gazella and hystrix indica, by setting the count to be equal to the highest count observed for the species.

```{r fix outliers}
maxI <- which(spp[,"Canis.lupus"]>40)
spp[maxI,"Canis.lupus"] <- max(spp[-maxI,"Canis.lupus"])
P.anal[maxI,`Canis lupus`:=max(spp[-maxI,"Canis.lupus"])]
abu_by_spp.ConiferForests[maxI,`Canis lupus`:=max(spp[-maxI,"Canis.lupus"])]

maxI <- which(spp[,"Gazella.gazella"]>35)
spp[maxI,"Gazella.gazella"] <- max(spp[-maxI,"Gazella.gazella"])
P.anal[maxI,`Gazella gazella`:=max(spp[-maxI,"Gazella.gazella"])]
abu_by_spp.ConiferForests[maxI,`Gazella gazella`:=max(spp[-maxI,"Gazella.gazella"])]

maxI <- which(spp[,"Hystrix.indica"]>30)
spp[maxI,"Hystrix.indica"] <- max(spp[-maxI,"Hystrix.indica"])
P.anal[maxI,`Hystrix indica`:=max(spp[-maxI,"Hystrix.indica"])]
abu_by_spp.ConiferForests[maxI,`Hystrix indica`:=max(spp[-maxI,"Hystrix.indica"])]

try(plot(spp ~ P.anal$Subunit, overall.main="raw abundances", transformation = "no"))
```

```{r data explore, fig.height = 8, fig.width = 10}
kable(summary(P.anal[,.(richness, abundance, Subunit, rescaled_Time.Diff, cosinus_Monitoring.Time.Diff, sinus_Monitoring.Time.Diff, Site, Transect, Transect_with_date, Distance_rescaled)]))
pairs(P.anal[,lapply(X = .SD,FUN = as.numeric),.SDcols=c("Subunit","sinus_Monitoring.Time.Diff", "cosinus_Monitoring.Time.Diff", "rescaled_Time.Diff","Site","Transect", "Transect_with_date", "Distance_rescaled")])
kable(cor(P.anal[,lapply(X = .SD,FUN = as.numeric),.SDcols=c("Subunit","sinus_Monitoring.Time.Diff", "cosinus_Monitoring.Time.Diff", "rescaled_Time.Diff","Site","Transect", "Transect_with_date", "Distance_rescaled")]))
meanvar.plot(spp, xlab = "mean abundance of a given species across sites", ylab = "variance of the abundance of a given species across sites")
```


#### Community analysis using package MVabund

Table of abundances per species by camera ID

```{r print table with rare species}
abundance_and_cameras <- as.data.table(abu_by_spp.ConiferForests[,c(5,27:39)])
dim(abundance_and_cameras)
canisau <- abundance_and_cameras[,2] > 0
sum(canisau)
canislup <- abundance_and_cameras[,3] > 0
sum(canislup)
capra <- abundance_and_cameras[,4] > 0
sum(capra)
Equus <- abundance_and_cameras[,5] > 0
sum(Equus)
Gazellador <- abundance_and_cameras[,6] > 0
sum(Gazellador)
Gazellagaz <- abundance_and_cameras[,7] > 0
sum(Gazellagaz)
Hyaena <- abundance_and_cameras[,8] > 0
sum(Hyaena)
Hystrix <- abundance_and_cameras[,9] > 0
sum(Hystrix)
Lepus <- abundance_and_cameras[,10] > 0
sum(Lepus)
Meles <- abundance_and_cameras[,11] > 0
sum(Meles)
Oryx <- abundance_and_cameras[,12] > 0
sum(Oryx)
Sus <- abundance_and_cameras[,13] > 0
sum(Sus)
Vulpus <- abundance_and_cameras[,14] > 0
sum(Vulpus)
```

298 cameras. Species with a minimum of 10 cameras - Canis aureus, Canis lupus, Gazella gazella, Hyaena hyaena, Hystrix indica, Lepus capensis, Meles meles, Sus scorfa and Vulpus vulpus

```{r filtering out rare species}

spp_no_rare <- abu_by_spp.ConiferForests[,27:39]
print(colSums(spp_no_rare))
# filter out species with less than 20 individuals (rare species)
spp_no_rare <- spp_no_rare[,.SD,.SDcols = colSums(spp_no_rare)>=20]
#spp_no_rare <- spp_no_rare[,c(-2,-6)]
spp_no_rare <- mvabund(spp_no_rare)
print(colSums(spp_no_rare))

```


Species with 20 individuals or more - Canis aureus, Canis lupus, Gazella gazella, Hyaena hyaena, Hystrix indica, Lepus capensis, Meles meles, Sus scorfa and Vulpus vulpus

Left with 9 species

```{r community analysis MVabund, fig.height=8, fig.width=10}

env_data <- abu_by_spp.ConiferForests[,..col_names]

mva_m0.po <- manyglm(formula = spp_no_rare ~ rescaled_Time.Diff + cosinus_Monitoring.Time.Diff + sinus_Monitoring.Time.Diff + Subunit, family = "poisson", data = env_data)

mva_m0.nb <- manyglm(formula = spp_no_rare ~ rescaled_Time.Diff + cosinus_Monitoring.Time.Diff + sinus_Monitoring.Time.Diff + Subunit, family = "negative.binomial", data = env_data)

aic_dt <- data.table(nb = mva_m0.nb$aic, po=mva_m0.po$aic)
colMeans(aic_dt)
print("POISSON")
plot(mva_m0.po, which=1:3)
print("NEGATIVE BINOMIAL")
plot(mva_m0.nb, which=1:3)
```

prefer negative binomial because standardized residuals are smaller.

```{r community analysis MVabund 2}
# mva_m0 <- mva_m0.po
# mva_m1 <- manyglm(formula = spp_no_rare ~ Distance_rescaled * rescaled_Time.Diff + dunes*rescaled_Time.Diff + Distance_rescaled:dunes + cos_td_rad + sin_td_rad + site, family = "poisson", data = env_data)
# aic_dt$po1 <- mva_m1$aic
mva_m0 <- mva_m0.nb
mva_m1 <- manyglm(formula = spp_no_rare ~ rescaled_Time.Diff + cosinus_Monitoring.Time.Diff + sinus_Monitoring.Time.Diff, data = env_data)
aic_dt$nb1 <- mva_m1$aic
colMeans(aic_dt)

mva_m2 <- manyglm(formula = spp_no_rare ~ rescaled_Time.Diff + cosinus_Monitoring.Time.Diff + sinus_Monitoring.Time.Diff + Transect, data = env_data)
aic_dt$nb2 <- mva_m2$aic
colMeans(aic_dt)

mva_m3 <- manyglm(formula = spp_no_rare ~ rescaled_Time.Diff + cosinus_Monitoring.Time.Diff + sinus_Monitoring.Time.Diff + Site, data = env_data)
aic_dt$nb3 <- mva_m2$aic
colMeans(aic_dt)

mva_m4 <- manyglm(formula = spp_no_rare ~ rescaled_Time.Diff + cosinus_Monitoring.Time.Diff + sinus_Monitoring.Time.Diff + Subunit, data = env_data)
aic_dt$nb4 <- mva_m2$aic
colMeans(aic_dt)

mva_m4 <- manyglm(formula = spp_no_rare ~ rescaled_Time.Diff + cosinus_Monitoring.Time.Diff + sinus_Monitoring.Time.Diff + Subunit + Site, data = env_data)
aic_dt$nb5 <- mva_m2$aic
colMeans(aic_dt)

```

Transect, Site and Subunit improve the model. Will proceed with Subunit:


## Many GLM with Sub-unit - no interactions

```{r community analysis (site) MVabund model selection}
mva_m0.nb <- manyglm(formula = spp_no_rare ~ rescaled_Time.Diff + cosinus_Monitoring.Time.Diff + sinus_Monitoring.Time.Diff + Subunit, family = "negative.binomial", data = env_data)
drop1(mva_m0.nb)

```

Drop sine

```{r community analysis (site) MVabund model selection 2}
# mva_m2 <- manyglm(formula = spp_no_rare ~ Distance_rescaled + dunes*rescaled_Time.Diff + Distance_rescaled:dunes + cos_td_rad + sin_td_rad, family = "poisson", data = env_data)
mva_m2 <- manyglm(formula = spp_no_rare ~ rescaled_Time.Diff + cosinus_Monitoring.Time.Diff + Subunit, family = "negative.binomial", data = env_data)
drop1(mva_m2)
```

 Final model includes Sub-unit, cosin and time

```{r community analysis (site) MVabund model validation, fig.height=8, fig.width=10}
plot(mva_m2, which=1:3)
```

```{r community analysis (site) MVabund model interpretation}

# summ_m2 <- summary(mva_m2)
# saveRDS(summ_m2, "output/summ_model_forest_subunit1.RDS")
summ_m2 <- readRDS("output/summ_model_forest_subunit1.RDS")
# anov_m2.uni <- anova(mva_m2, p.uni = "adjusted")
# saveRDS(anov_m2.uni, "output/anova_forest_subunit1.RDS")
anov_m2.uni <- readRDS("output/anova_forest_subunit1.RDS")
print(summ_m2)
print(anov_m2.uni)

```

**Multivariate test shows that Sub-unit, cosin time of sampling and time affect species composition**
**However time variable is not significant for any of the species. Therefore re-run without time variable.**

```{r}
mva_m2 <- manyglm(formula = spp_no_rare ~ cosinus_Monitoring.Time.Diff + Subunit, family = "negative.binomial", data = env_data)
drop1(mva_m2)
```
```{r}
plot(mva_m2, which=1:3)
```
```{r}
# summ_m2 <- summary(mva_m2)
# saveRDS(summ_m2, "output/summ_model_forest_subunit2.RDS")
summ_m2 <- readRDS("output/summ_model_forest_subunit2.RDS")
# anov_m2.uni <- anova(mva_m2, p.uni = "adjusted")
# saveRDS(anov_m2.uni, "output/anova_forest_subunit2.RDS")
anov_m2.uni <- readRDS("output/anova_forest_subunit2.RDS")
print(summ_m2)
print(anov_m2.uni)
```

**Final model includes Sub-unit and cosin time of sampling**

```{r community analysis (site) MVabund model interpretation - plot coefficients, error=TRUE}
coefp <- merge(data.table(t(coef(mva_m2)/log(2)),keep.rownames=TRUE), data.table(t(anov_m2.uni$uni.p), keep.rownames = TRUE), by = "rn") # coefficients are exponentiated because the link function used for poisson is log -> above 1 is positive and below 1 is negative
# add total species abundance
coefp <- merge(coefp,as.data.table(colSums(spp_no_rare),keep.rownames = TRUE), by.x = "rn", by.y = "V1")

colnames(coefp) <- c("SciName","Intercept.coef","Cosin_time.coef","Galilee.coef","Judea.coef","Intercept.p","Cosin_time.p","Subunit.p","Species_abundance")

write.csv(coefp, "coefficients_Subunit_Planted_Conifer_Forest.csv")

```


### Canis lupus
remove Carmel observations because they are all zeros.

```{r community analysis MVabund model interpretation - single species models, error=TRUE}
Canis.lupus <- spp_no_rare[env_data$Subunit!="Carmel","Canis.lupus"]
glm.Canis.lupus <- glm.nb(formula = Canis.lupus ~ cosinus_Monitoring.Time.Diff  + Subunit, data = env_data[Subunit!="Carmel"])
summary(glm.Canis.lupus)

plot_model_effect(P.anal = env_data[Subunit!="Carmel"][, Canis.lupus:=as.numeric(Canis.lupus)], m = glm.Canis.lupus, eff2plot = "Subunit", plot_points=TRUE, plot_residuals=FALSE, export_plot=TRUE, ylabel = NULL, fontname = "Almoni ML v5 AAA", fontsize=22, pdf_width=160, outpath = "output/conifer_forest/", y_cutoff = 4)
```

Compute predicted average difference in C. lupus abundance between Judea and Galilee based on model:

```{r}
I_td_min <- which.min(P.anal$rescaled_Time.Diff)
I_td_max <- which.max(P.anal$rescaled_Time.Diff)
abund_galilee_judea_canlup <- predict(object = glm.Canis.lupus,
                              newdata = list(rescaled_Time.Diff=rep(mean(c(P.anal$rescaled_Time.Diff[I_td_min],
                                                                           P.anal$rescaled_Time.Diff[I_td_max])),2),
                                             Subunit = c("Galilee","Judea"),
                                             cosinus_Monitoring.Time.Diff = rep(cos(mean(P.anal$Monitoring.Time.Diff_numeric)),2)),
                              type = "response")
print(paste("C. lupus abundance in Galilee is higher than Judea by",
      abs(diff(abund_galilee_judea_canlup)/abund_galilee_judea_canlup[2]*100),"%"))
```

### Gazella gazella

```{r}
Gazella.gazella <- spp_no_rare[,"Gazella.gazella"]
glm.Gazella.gazella <- glm.nb(formula = Gazella.gazella ~ cosinus_Monitoring.Time.Diff  + Subunit, data = env_data)
coef(glm.Gazella.gazella)
coef(mva_m2)
print(sum(Gazella.gazella))

plot_model_effect(P.anal = env_data, m = glm.Gazella.gazella, eff2plot = "Subunit", plot_points=TRUE, plot_residuals=FALSE, export_plot=TRUE, ylabel = NULL, fontname = "Almoni ML v5 AAA", fontsize=22, pdf_width=160, outpath = "output/conifer_forest/")

#Compare spatial trends among subunits
pairwise_subunit <- emmeans(object = glm.Gazella.gazella, ~ Subunit)
print(pairwise_subunit)
test_results_subunit <- test(pairs(pairwise_subunit, by=NULL), adjust="fdr")
print(test_results_subunit)
```

Compute predicted average difference in G. gazella abundance among all subunits based on model:

```{r}
abund_gazgaz <- predict(object = glm.Gazella.gazella,
                              newdata = list(rescaled_Time.Diff=rep(mean(c(P.anal$rescaled_Time.Diff[I_td_min],
                                                                           P.anal$rescaled_Time.Diff[I_td_max])),3),
                                             Subunit = c("Judea","Galilee","Carmel"),
                                             cosinus_Monitoring.Time.Diff = rep(cos(mean(P.anal$Monitoring.Time.Diff_numeric)),3)),
                              type = "response")
print(paste("G. gazella abundance in Judea is higher than Carmel by",
      abs(diff(abund_gazgaz[c(1,3)])/abund_gazgaz[3]*100),"%"))

print(paste("G. gazella abundance in Judea is higher than Galilee by",
      abs(diff(abund_gazgaz[c(1,2)])/abund_gazgaz[2]*100),"%"))

print(paste("G. gazella abundance in Galilee is higher than Carmel by",
      abs(diff(abund_gazgaz[c(2,3)])/abund_gazgaz[3]*100),"%"))
```

### Hystrix indica

```{r}
Hystrix.indica <- spp_no_rare[,"Hystrix.indica"]
glm.Hystrix.indica <- glm.nb(formula = Hystrix.indica ~ cosinus_Monitoring.Time.Diff  + Subunit, data = env_data)
coef(glm.Hystrix.indica)
coef(mva_m2)

plot_model_effect(P.anal = env_data, m = glm.Hystrix.indica, eff2plot = "Subunit", plot_points=TRUE, plot_residuals=FALSE, export_plot=TRUE, ylabel = NULL, fontname = "Almoni ML v5 AAA", fontsize=22, pdf_width=160, outpath = "output/conifer_forest/")

#Compare spatial trends among subunits
pairwise_subunit <- emmeans(object = glm.Hystrix.indica, ~ Subunit)
print(pairwise_subunit)
test_results_subunit <- test(pairs(pairwise_subunit, by=NULL), adjust="fdr")
print(test_results_subunit)
```

Compute predicted average difference in H. indica abundance among all subunits based on model:

```{r}
abund_hysind <- predict(object = glm.Hystrix.indica,
                              newdata = list(rescaled_Time.Diff=rep(mean(c(P.anal$rescaled_Time.Diff[I_td_min],
                                                                           P.anal$rescaled_Time.Diff[I_td_max])),3),
                                             Subunit = c("Judea","Carmel","Galilee"),
                                             cosinus_Monitoring.Time.Diff = rep(cos(mean(P.anal$Monitoring.Time.Diff_numeric)),3)),
                              type = "response")
print(paste("H. indica abundance in Judea is higher than Carmel by",
      abs(diff(abund_hysind[c(1,2)])/abund_hysind[2]*100),"%"))

print(paste("H. indica abundance in Judea is higher than Galilee by",
      abs(diff(abund_hysind[c(1,3)])/abund_hysind[3]*100),"%"))

print(paste("H. indica abundance in Carmel is higher than Galilee by",
      abs(diff(abund_hysind[c(2,3)])/abund_hysind[3]*100),"%"))
```

### Lepus cepensis
remove Galilee observations because they are all zeros.

```{r}
Lepus.capensis <- spp_no_rare[env_data$Subunit!="Galilee","Lepus.capensis"]
glm.Lepus.capensis <- glm.nb(formula = Lepus.capensis ~ cosinus_Monitoring.Time.Diff  + Subunit, data = env_data[Subunit!="Galilee"])
summary(glm.Lepus.capensis)
```

After removing Galilee, subunit is not a significant factor on Lepus capensis abundance.

### Sus scrofa
although significant, exclude model because appears again later in model with subunit*time interaction.

```{r, eval=FALSE}
Sus.scrofa <- spp_no_rare[,"Sus.scrofa"]
glm.Sus.scrofa <- glm.nb(formula = Sus.scrofa ~ cosinus_Monitoring.Time.Diff  + Subunit, data = env_data)
coef(glm.Sus.scrofa)
coef(mva_m2)

plot_model_effect(P.anal = env_data, m = glm.Sus.scrofa, eff2plot = "Subunit", plot_points=TRUE, plot_residuals=FALSE, export_plot=TRUE, ylabel = NULL, fontname = "Almoni ML v5 AAA", fontsize=22, pdf_width=160, outpath = "output/conifer_forest/")

#Compare spatial trends among subunits
pairwise_subunit <- emmeans(object = glm.Sus.scrofa, ~ Subunit)
print(pairwise_subunit)
test_results_subunit <- test(pairs(pairwise_subunit, by=NULL), adjust="fdr")
print(test_results_subunit)

```


## Many GLM with interaction between subunit and time

```{r community analysis (Subunit) MVabund model selection}

mva_m0.nb <- manyglm(formula = spp_no_rare ~ Subunit * rescaled_Time.Diff + cosinus_Monitoring.Time.Diff + sinus_Monitoring.Time.Diff, family = "negative.binomial", data = env_data)
drop1(mva_m0.nb)

```

drop sine

```{r}
mva_m0.nb <- manyglm(formula = spp_no_rare ~ Subunit * rescaled_Time.Diff + cosinus_Monitoring.Time.Diff, family = "negative.binomial", data = env_data)
drop1(mva_m0.nb)
```

Final model includes cosin time, and the interaction between subunit and time 

```{r community analysis (Subunit) MVabund model interpretation}
# summ_m2 <- summary(mva_m0.nb)
# saveRDS(summ_m2, "output/summ_model_forest_interact.RDS")
summ_m2 <- readRDS("output/summ_model_forest_interact.RDS")
print(summ_m2)
# anov_m2.uni <- anova(mva_m0.nb, p.uni = "adjusted")
# saveRDS(anov_m2.uni, "output/anova_forest_interact.RDS")
anov_m2.uni <- readRDS("output/anova_forest_interact.RDS")
print(anov_m2.uni)
```

**Multivariate test shows that subunit x time and cosin time affect species composition**

```{r community analysis (subunit) MVabund model interpretation - plot coefficients, error=TRUE}
coefp <- merge(data.table(t(coef(mva_m0.nb)/log(2)),keep.rownames=TRUE), data.table(t(anov_m2.uni$uni.p), keep.rownames = TRUE), by = "rn") # coefficients are exponentiated because the link function used for poisson is log -> above 1 is positive and below 1 is negative
# add total species abundance
coefp <- merge(coefp,as.data.table(colSums(spp_no_rare),keep.rownames = TRUE), by.x = "rn", by.y = "V1")

colnames(coefp) <- c("SciName","Intercept.coef","Galilee.coef","Judea.coef","time_diff","Cosin_time.coef","Galilee_time.coef","Judea_time.coef","Intercept.p","Subunit.p","Time_diff.p","Cosin.p","Subunit_time.p","Species_abundance")

write.csv(coefp, "coefficients_Subunit_interaction_Planted_Conifer_Forest.csv")
```


### Sus scrofa

```{r, error=TRUE}
Sus.scrofa <- spp_no_rare[,"Sus.scrofa"]
glm.Sus.scrofa <- glm.nb(Sus.scrofa ~ Subunit * rescaled_Time.Diff + cosinus_Monitoring.Time.Diff, data = env_data, maxit = 999)
coef(glm.Sus.scrofa)
coef(mva_m0.nb)

interact_plot(glm.Sus.scrofa, pred = rescaled_Time.Diff, modx = Subunit, modx.values = NULL, partial.residuals = F, jitter = c(0.2), point.size = 3, rug = F, point.shape = F, interval = T, colors = "Dark2")

plot_model_interaction(P.anal = env_data, m = glm.Sus.scrofa, eff2plot = "rescaled_Time.Diff", modvar2plot = "Subunit", plot_points=FALSE, plot_residuals=FALSE, export_plot=TRUE, ylabel = NULL, fontname = "Almoni ML v5 AAA", fontsize=22, pdf_width=160, outpath = "output/conifer_forest/", legend_position = "bottom")
```

Test whether temporal trend is different from zero

```{r, error=TRUE}
#Test whether temporal trend is different from zero
mm_rescaled_subunit_time <- emtrends(glm.Sus.scrofa, specs = "Subunit", var = "rescaled_Time.Diff", type = "response")
print(mm_rescaled_subunit_time)
test_subunit_time <- test(mm_rescaled_subunit_time, null = 0, adjust="fdr")
print(test_subunit_time)
```

Compute predicted decrease in S. scrofa across monitoring time span based on model:

```{r increase in S. scrofa in Judea}
print(paste("start date:", P.anal$Start.date[I_td_min]))
print(paste("end date:", P.anal$Start.date[I_td_max]))

abund_start_end <- predict(object = glm.Sus.scrofa, newdata = list(rescaled_Time.Diff=c(P.anal$rescaled_Time.Diff[I_td_min], P.anal$rescaled_Time.Diff[I_td_max]), Subunit = c("Judea","Judea"), cosinus_Monitoring.Time.Diff = rep(cos(mean(P.anal$Monitoring.Time.Diff_numeric)),2)), type = "response")

print(paste("Increase from",abund_start_end[1],"to", abund_start_end[2], "which is",
            diff(abund_start_end)/abund_start_end[1]*100,"%"))
```

Test whether spatial differences in subunit are different from zero

```{r, error=TRUE}
#Test whether spatial differences in subunit are different from zero
EMM <- emmeans(object = glm.Sus.scrofa, ~Subunit*rescaled_Time.Diff)
test_results_subunit <- test(pairs(EMM, by="rescaled_Time.Diff"), by=NULL, adjust="fdr")
print(test_results_subunit)
```

Compute predicted average difference in S. scrofa abundance between Judea-Carmel and Judea-Galilee based on model:

```{r}

abund_susscr <- predict(object = glm.Sus.scrofa,
                              newdata = list(rescaled_Time.Diff=rep(mean(c(P.anal$rescaled_Time.Diff[I_td_min],
                                                                           P.anal$rescaled_Time.Diff[I_td_max])),3),
                                             Subunit = c("Judea","Carmel","Galilee"),
                                             cosinus_Monitoring.Time.Diff = rep(cos(mean(P.anal$Monitoring.Time.Diff_numeric)),3)),
                              type = "response")
print(paste("S. scrofa abundance in Carmel is higher than Judea by",
      abs(diff(abund_susscr[c(1,2)])/abund_susscr[1]*100),"%"))

print(paste("S. scrofa abundance in Galilee is higher than Judea by",
      abs(diff(abund_susscr[c(1,3)])/abund_susscr[1]*100),"%"))
```

### Canis aureus

```{r, error=TRUE}
Canis.aureus <- spp_no_rare[,"Canis.aureus"]
glm.Canis.aureus <- glm.nb(Canis.aureus ~ Subunit * rescaled_Time.Diff + cosinus_Monitoring.Time.Diff, data = env_data, maxit = 999)
coef(glm.Canis.aureus)
coef(mva_m0.nb)

# interact_plot(glm.Canis.aureus, pred = rescaled_Time.Diff, modx = Subunit, modx.values = NULL, partial.residuals = F, jitter = c(0.2), point.size = 3, rug = F, point.shape = F, interval = T, colors = "Dark2")

plot_model_interaction(P.anal = env_data, m = glm.Canis.aureus, eff2plot = "rescaled_Time.Diff", modvar2plot = "Subunit", plot_points=FALSE, plot_residuals=FALSE, export_plot=TRUE, ylabel = NULL, fontname = "Almoni ML v5 AAA", fontsize=22, pdf_width=160, outpath = "output/conifer_forest/", legend_position = "bottom")
```

Test whether temporal trend is different from zero

```{r, error=TRUE}
#Test whether temporal trend is different from zero
print("Canis aureus")
mm_rescaled_subunit_time <- emtrends(glm.Canis.aureus, specs = "Subunit", var = "rescaled_Time.Diff", type = "response")
print(mm_rescaled_subunit_time)
test_subunit_time <- test(mm_rescaled_subunit_time, null = 0, adjust="fdr")
print(test_subunit_time)
```

Compute predicted decrease in C. aureus across monitoring time span based on model:

```{r decrease in C. aureus in Carmel}
print(paste("start date:", P.anal$Start.date[I_td_min]))
print(paste("end date:", P.anal$Start.date[I_td_max]))

abund_start_end <- predict(object = glm.Canis.aureus, newdata = list(rescaled_Time.Diff=c(P.anal$rescaled_Time.Diff[I_td_min], P.anal$rescaled_Time.Diff[I_td_max]), Subunit = c("Carmel","Carmel"), cosinus_Monitoring.Time.Diff = rep(cos(mean(P.anal$Monitoring.Time.Diff_numeric)),2)), type = "response")

print(paste("Decrease from",abund_start_end[1],"to", abund_start_end[2], "which is",
            diff(abund_start_end)/abund_start_end[1]*100,"%"))
```


Test whether spatial differences in subunit are different from zero

```{r, error=TRUE}
#Test whether spatial differences in subunit are different from zero
EMM <- emmeans(object = glm.Canis.aureus, ~Subunit*rescaled_Time.Diff)
test_results_subunit <- test(pairs(EMM, by="rescaled_Time.Diff"), by=NULL, adjust="fdr")
print(test_results_subunit)
```
Compute predicted average difference in C. aureus abundance between Carmel and Judea based on model:

```{r}

abund_carmel_judea_canaur <- predict(object = glm.Canis.aureus,
                              newdata = list(rescaled_Time.Diff=rep(mean(c(P.anal$rescaled_Time.Diff[I_td_min],
                                                                           P.anal$rescaled_Time.Diff[I_td_max])),2),
                                             Subunit = c("Carmel","Judea"),
                                             cosinus_Monitoring.Time.Diff = rep(cos(mean(P.anal$Monitoring.Time.Diff_numeric)),2)),
                              type = "response")
print(paste("C. aureus abundance in Carmel is higher than Judea by",
      abs(diff(abund_carmel_judea_canaur)/abund_carmel_judea_canaur[2]*100),"%"))
```

n = 551 (total number of cameras)

total abundance = 5613 (9 species), total abundance without rare species is the same

### Session information

```{r}
session_info()
```

