---
title: "Mammal data - Mediterranean Maquis - with subunit interactions"
author: "Ron Chen & Talia Gabay"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)
#knitr::opts_chunk$set(dev = "cairo_pdf", dev.args=list(cairo_pdf = list(family='Arial')))

#Load packages
library(devtools)
library(rlang)
library(ggplot2)
library(data.table)
library(dplyr)
library(mvabund)
library(kableExtra)
library(lme4)
library(MASS)
library(vegan)
library(jtools)
library(interactions)
library(car)
library(ecoCopula)
library(chron)
library(data.table)
library(readxl)
library(performance)
library(Cairo)
library(emmeans)
library(extrafont)

#Load functions
source("functions/plot_alpha_diversity_cycle_num.R")
source("functions/plot_model_effect_mammals_Maquis_and_Forest.R")
source("functions/plot_interaction_mammals_Maquis_and_Forest.R")
source_url("https://raw.githubusercontent.com/ronchen1/ron_functions/d2d0bab4dc65f29e53d316ae4098d0469711905b/R/multiplot.R", sha1 = "eef55df1963ba201f57cf44513f566298756fd4f")

```



```{r load data, include=FALSE}
# Abundance table - all data
all.mammal.abundances <- read.csv("data/mammals.abundance.matrix.TG.csv", header = TRUE, sep =",") %>% as.data.table()


#Species names
sci_names <- readxl::read_excel(path = 'data/Scientific.names.mammals.xlsx') %>% as.data.table()
#Make species list
sci_names_list <- as.character(sort(unique(sci_names$SciName)))
#Change species column names in abundance table 
colnames(all.mammal.abundances)[27:40] <- sci_names_list 
#colnames(all.mammal.abundances)[25:38] <- sci_names_list   
   
all.mammal.abundances <- all.mammal.abundances[,-30]
 
  P_byplot <- copy(all.mammal.abundances)
  P_byplot[,`:=`(richness = apply(.SD,1,(\(x) sum(x>0))),
                 abundance = apply(.SD,1,sum),
                 gma=apply(.SD,1,(\(x) exp( sum(log(x[x>0])) / sum(x>0) )))),
           .SDcols = sci_names_list[-4]]

```

```{r load using function}

P <- P_byplot[grepl("Mediterranean Maquis",Unit),][order(Cycle_number,Site)][,Site:=factor(Site)]
```

## Mediterranean Maquis - Richness by camera

### Monitoring started in 2014 (5 monitoring cycles), Each site has two types of plots - far or near settlements
### There are 15 sites total, each site has 5 sampling points 

Let's first validate the data:

```{r data validation}
#include rare species in analysis
P.anal <- copy(P) # set a fixed variable name for analysis, if want to switch between data WITH rare #species and data WITHOUT rare species then only change once here
#Validate factors and levels
P.anal$Settlements <- as.factor(P.anal$Settlements)
print("Settlements has 2 levels")
print(levels(P.anal$Settlements))
P.anal$Subunit <- as.factor(P.anal$Subunit)
print("Subunit has 3 levels")
print(levels(P.anal$Subunit))
P.anal$Site <- as.factor(P.anal$Site)
print("Site has 15 levels")
print(levels(P.anal$Site))
P.anal$Transect <- as.factor(P.anal$Transect)
print("Transect has 30 levels")
print(levels(P.anal$Transect))
P.anal$Transect_with_date <- as.factor(P.anal$Transect_with_date)
print("Transect with date has 150 levels")
print(levels(P.anal$Transect_with_date))

```
```{r richness data explore, fig.height = 8, fig.width = 10}
print("RICHNESS WITH RARE SPECIES")
plot_alpha_diversity(P, x_val = "Subunit", y_val = "richness", ylab_val = "richness", xlab_val = "Subunit", fill_val = "Subunit")

col_names <- c("Cycle_number",  "Deployment.id_new",	"Unit",	"Subunit",	"Site", "Settlements",	"Distance_rescaled", "Lon",	"Lat", "year", "rescaled_Time.Diff", "sinus_Monitoring.Time.Diff", "cosinus_Monitoring.Time.Diff", "Transect", "Transect_with_date")

abu_by_spp.Maquis <- all.mammal.abundances[grepl("Mediterranean Maquis",Unit),]
spp <- abu_by_spp.Maquis[,c(27:39)]

# filter out species with zero counts 
print(colSums(spp))
spp <- spp[,.SD,.SDcols = colSums(spp)>0]
spp <- mvabund(spp)

plot(spp ~ P.anal$Subunit, overall.main="raw abundances", transformation = "no")

my_cols <- c("blue", "darkgreen", "red")
dotchart(P.anal$richness, ylab = "Subunit",
         groups = P.anal$Subunit, gcolor = my_cols,
         color = my_cols[P.anal$Subunit],
         cex = 0.9,  pch = 1, xlab = "richness")

kable(summary(P.anal[,.(richness, abundance, Subunit, rescaled_Time.Diff, cosinus_Monitoring.Time.Diff, sinus_Monitoring.Time.Diff, Site, Transect, Transect_with_date, Distance_rescaled)]))
pairs(P.anal[,lapply(X = .SD,FUN = as.numeric),.SDcols=c("Subunit","sinus_Monitoring.Time.Diff", "cosinus_Monitoring.Time.Diff", "rescaled_Time.Diff","Site","Transect", "Transect_with_date", "Distance_rescaled")])
kable(cor(P.anal[,lapply(X = .SD,FUN = as.numeric),.SDcols=c("Subunit","sinus_Monitoring.Time.Diff", "cosinus_Monitoring.Time.Diff", "rescaled_Time.Diff","Site","Transect", "Transect_with_date", "Distance_rescaled")]))
meanvar.plot(spp, xlab = "mean abundance of a given species across sites", ylab = "variance of the abundance of a given species across sites")

```

## Community analysis using package MVabund


Table of abundances per species by camera ID

```{r print table with rare species}
abundance_and_cameras <- as.data.table(abu_by_spp.Maquis[,c(5,27:39)])
dim(abundance_and_cameras)
canisau <- abundance_and_cameras[,2] > 0
sum(canisau)
canislup <- abundance_and_cameras[,3] > 0
sum(canislup)
capra <- abundance_and_cameras[,4] > 0
sum(capra)
Equus <- abundance_and_cameras[,5] > 0
sum(Equus)
Gazellador <- abundance_and_cameras[,6] > 0
sum(Gazellador)
Gazellagaz <- abundance_and_cameras[,7] > 0
sum(Gazellagaz)
Hyaena <- abundance_and_cameras[,8] > 0
sum(Hyaena)
Hystrix <- abundance_and_cameras[,9] > 0
sum(Hystrix)
Lepus <- abundance_and_cameras[,10] > 0
sum(Lepus)
Meles <- abundance_and_cameras[,11] > 0
sum(Meles)
Oryx <- abundance_and_cameras[,12] > 0
sum(Oryx)
Sus <- abundance_and_cameras[,13] > 0
sum(Sus)
Vulpus <- abundance_and_cameras[,14] > 0
sum(Vulpus)
```

1117 cameras. Species with a minimum of 20 cameras - Canis aureus, Canis lupus, Gazella gazella, Hyaena hyaena, Hystrix indica, Meles meles, Sus scorfa and Vulpus vulpus

```{r filtering out rare species}

spp_no_rare <- abu_by_spp.Maquis[,27:39]
print(colSums(spp_no_rare))
# filter out species with less than 20 individuals (rare species)
spp_no_rare <- spp_no_rare[,.SD,.SDcols = colSums(spp_no_rare)>=20]
#spp_no_rare <- spp_no_rare[,c(-2,-6)]
spp_no_rare <- mvabund(spp_no_rare)
print(colSums(spp_no_rare))

#Sus scorfa abundance across subunit, sites and projectid
my_cols <- c("blue", "darkgreen", "red")
dotchart(P.anal$`Sus scrofa`, ylab = "Subunit",
         groups = P.anal$Subunit, gcolor = my_cols,
         color = my_cols[P.anal$Subunit],
         cex = 0.9,  pch = 1, xlab = "Abundance")

P.anal$Project.ID <- as.factor(P.anal$Project.ID)
my_cols <- c("blue", "darkgreen", "red", "purple", "darkorange")
dotchart(P.anal$`Sus scrofa`, ylab = "Project.ID",
         groups = P.anal$Project.ID, gcolor = my_cols,
         color = my_cols[P.anal$Project.ID],
         cex = 0.9,  pch = 1, xlab = "Abundance")

fun_color_range <- colorRampPalette(c("#1b98e0", "red"))
my_cols <- fun_color_range(15)    
dotchart(P.anal$`Sus scrofa`, ylab = "Site",
         groups = P.anal$Site, gcolor = my_cols,
         color = my_cols[P.anal$Site],
         cex = 0.9,  pch = 1, xlab = "Abundance")

```

### Species with 20 individuals or more - Canis aureus, Canis lupus, Gazella gazella, Hyaena hyaena, Hystrix indica, Meles meles, Sus scorfa and Vulpus vulpus

### Left with 8 species

```{r community analysis MVabund, fig.height=8, fig.width=10}
env_data <- abu_by_spp.Maquis[,..col_names]
```


# Many GLM with subunit interactions 

```{r community analysis (Subunit interactions) MVabund model selection}
mva_m0.nb <- manyglm(formula = spp_no_rare ~ Distance_rescaled * rescaled_Time.Diff + Subunit * Distance_rescaled + Subunit * rescaled_Time.Diff + cosinus_Monitoring.Time.Diff + sinus_Monitoring.Time.Diff, family = "negative.binomial", data = env_data)
drop1(mva_m0.nb)

```

```{r community analysis (Subunit interactions) MVabund model selection 2}
mva_m1 <- manyglm(formula = spp_no_rare ~ Distance_rescaled * rescaled_Time.Diff + Subunit * Distance_rescaled + Subunit * rescaled_Time.Diff + cosinus_Monitoring.Time.Diff, family = "negative.binomial", data = env_data)
drop1(mva_m1)

```
```{r community analysis (Subunit interactions) MVabund model selection 3}
mva_m1 <- manyglm(formula = spp_no_rare ~ Subunit * Distance_rescaled + Subunit * rescaled_Time.Diff + cosinus_Monitoring.Time.Diff, family = "negative.binomial", data = env_data)
drop1(mva_m1)

```

### Model includes cosin time, interaction between subunit and time and subunit and distance

```{r community analysis (Subunit interactions) MVabund model validation, fig.height=8, fig.width=10}
plot(mva_m1, which=1:3)
```


```{r community analysis (Subunit interactions) MVabund model interpretation}

# summ_m1 <- summary(mva_m1)
# saveRDS(summ_m1, "output/summ_maquis_model_with_subunit_interaction.RDS")
summ_m1 <- readRDS("output/summ_maquis_model_with_subunit_interaction.RDS")
print(summ_m1)
# anov_m1.uni <- anova(mva_m1, p.uni = "adjusted")
# saveRDS(anov_m1.uni, "output/anova_maquis_model_with_subunit_interaction.RDS")
anov_m1.uni <- readRDS("output/anova_maquis_model_with_subunit_interaction.RDS")
print(anov_m1.uni)
```

```{r community analysis (subunit interactions) MVabund model interpretation - plot coefficients}

coefp <- merge(data.table(t(coef(mva_m1)/log(2)),keep.rownames=TRUE), data.table(t(anov_m1.uni$uni.p), keep.rownames = TRUE), by = "rn") # coefficients are exponentiated because the link function used for poisson is log -> above 1 is positive and below 1 is negative
# add total species abundance
coefp <- merge(coefp,as.data.table(colSums(spp_no_rare),keep.rownames = TRUE), by.x = "rn", by.y = "V1")

colnames(coefp) <- c("SciName","Intercept.coef","Galilee.coef","Judea.coef","Distance.coef","Time_diff.coef","cosin_time.coef","Galilee_distance.coef","Judea_distance.coef","Galilee_time.coef","Judea_time.coef","Intercept.p","Subunit.p","Distance.p","Time_diff.p","Cosin_time.p","Subunit_distance.p","Subunit_time.p","Species_abundance")

write.csv(coefp, "coefficients_Maquis_with_subunit_interactions.csv")


```

### Canis aureus

```{r community analysis (subunit interactions) - per species models Canis aureus}
#####Canis aureus####

Canis.aureus <- spp_no_rare[,"Canis.aureus"]
glm.Canis.aureus <- glm.nb(formula = Canis.aureus ~ Subunit * Distance_rescaled + Subunit * rescaled_Time.Diff + cosinus_Monitoring.Time.Diff, data = env_data)
coef(glm.Canis.aureus)
coef(mva_m1)

plot_model_interaction(P.anal = env_data, m = glm.Canis.aureus, eff2plot = "Distance_rescaled", modvar2plot = "Subunit", plot_points=FALSE, plot_residuals=FALSE, export_plot=TRUE, ylabel = NULL, fontname = "Almoni ML v5 AAA",fontsize=22, pdf_width=160, outpath = "output/maquis/", legend_position = "bottom")

plot_model_interaction(P.anal = env_data, m = glm.Canis.aureus, eff2plot = "rescaled_Time.Diff", modvar2plot = "Subunit", plot_points=FALSE, plot_residuals=FALSE, export_plot=TRUE, ylabel = NULL, fontname = "Almoni ML v5 AAA",fontsize=22, pdf_width=160, outpath = "output/maquis/", legend_position = "bottom")

#Test whether temporal trend is different from zero
print("Test whether temporal trend is different from zero")
mm_rescaled_time <- emtrends(glm.Canis.aureus, specs = "Subunit", var = "rescaled_Time.Diff", type = "response")
print(mm_rescaled_time)
test_results_mm_rescaled_time <- test(mm_rescaled_time, null = 0, adjust = "fdr")
print(test_results_mm_rescaled_time)

#Test whether spatial trend is different from zero
print("Test whether spatial trend is different from zero")
mm_rescaled_distance <- emtrends(glm.Canis.aureus, specs = "Subunit", var = "Distance_rescaled", type = "response")
print(mm_rescaled_distance)
test_results_mm_distance <- test(mm_rescaled_distance, null = 0, adjust = "fdr")
print(test_results_mm_distance)

#Compare spatial trends among subunits
print("Compare spatial trends among subunits")
pairwise_distance <- emmeans(object = glm.Canis.aureus, ~ Subunit*Distance_rescaled)
print(pairwise_distance)
test_results_distance <- test(pairs(pairwise_distance, by="Distance_rescaled"), by=NULL, adjust="fdr")
print(test_results_distance)
```

### Hyaena hyaena

```{r community analysis (subunit interactions) - per species Hyaena hyaena}
#####Hyaena hyaena####

Hyaena.hyaena <- spp_no_rare[,"Hyaena.hyaena"]
glm.Hyaena.hyaena <- glm.nb(formula = Hyaena.hyaena ~ Subunit * Distance_rescaled + Subunit * rescaled_Time.Diff + cosinus_Monitoring.Time.Diff, data = env_data)
coef(glm.Hyaena.hyaena)
coef(mva_m1)

plot_model_interaction(P.anal = env_data, m = glm.Hyaena.hyaena, eff2plot = "Distance_rescaled", modvar2plot = "Subunit", plot_points=FALSE, plot_residuals=FALSE, export_plot=TRUE, ylabel = NULL, fontname = "Almoni ML v5 AAA", fontsize=22, pdf_width=160, outpath = "output/maquis/", legend_position = "bottom")

#Test whether spatial trend is different from zero
print("Test whether spatial trend is different from zero")
mm_rescaled_distance <- emtrends(glm.Hyaena.hyaena, specs = "Subunit", var = "Distance_rescaled", type = "response")
print(mm_rescaled_distance)
test_results_mm_distance <- test(mm_rescaled_distance, null = 0, adjust = "fdr")
print(test_results_mm_distance)

#Compare spatial trends among subunits
print("Compare spatial trends among subunits")
pairwise_distance <- emmeans(object = glm.Hyaena.hyaena, ~ Subunit*Distance_rescaled)
print(pairwise_distance)
test_results_distance <- test(pairs(pairwise_distance, by="Distance_rescaled"), by=NULL, adjust="fdr")
print(test_results_distance)

```


# Hystrix indica

```{r community analysis (subunit interactions) - per species models - Hystrix indica1}
#####Hystrix indica####

Hystrix.indica <- spp_no_rare[,"Hystrix.indica"]
glm.Hystrix.indica <- glm.nb(formula = Hystrix.indica ~ Subunit * Distance_rescaled + Subunit * rescaled_Time.Diff + cosinus_Monitoring.Time.Diff, data = env_data)
summary(glm.Hystrix.indica)
```

**Interaction of subunit with time is not significant - will be dropped from H. indica model**

```{r community analysis (subunit interactions) - per species models - Hystrix indica2}
glm.Hystrix.indica <- glm.nb(formula = Hystrix.indica ~ Subunit * Distance_rescaled + rescaled_Time.Diff + cosinus_Monitoring.Time.Diff, data = env_data)
summary(glm.Hystrix.indica)
coef(glm.Hystrix.indica)
coef(mva_m1)

plot_model_interaction(P.anal = env_data, m = glm.Hystrix.indica, eff2plot = "Distance_rescaled", modvar2plot = "Subunit", plot_points=FALSE, plot_residuals=FALSE, export_plot=TRUE, ylabel = NULL, fontname = "Almoni ML v5 AAA",fontsize=22, pdf_width=160, outpath = "output/maquis/", legend_position = "bottom")


#Test whether spatial trend is different from zero
print("Test whether spatial trend is different from zero")
mm_rescaled_distance <- emtrends(glm.Hystrix.indica, specs = "Subunit", var = "Distance_rescaled", type = "response")
print(mm_rescaled_distance)
test_results_distance <- test(mm_rescaled_distance, null = 0, adjust="fdr")
print(test_results_distance)

plot_model_effect(P.anal = env_data, m = glm.Hystrix.indica, eff2plot = "rescaled_Time.Diff", plot_points=FALSE, plot_residuals=FALSE, export_plot=TRUE, ylabel = NULL, fontname = "Almoni ML v5 AAA", fontsize=22, pdf_width=160, outpath = "output/maquis/")
```

### Meles meles

```{r community analysis (subunit interactions) - per species models Meles meles}
#####Meles meles####

Meles.meles <- spp_no_rare[,"Meles.meles"]
glm.Meles.meles <- glm.nb(formula = Meles.meles ~ Subunit * Distance_rescaled + Subunit * rescaled_Time.Diff + cosinus_Monitoring.Time.Diff, data = env_data)
coef(glm.Meles.meles)
coef(mva_m1)

plot_model_interaction(P.anal = env_data, m = glm.Meles.meles, eff2plot = "Distance_rescaled", modvar2plot = "Subunit", plot_points=FALSE, plot_residuals=FALSE, export_plot=TRUE, ylabel = NULL, fontname = "Almoni ML v5 AAA",fontsize=22, pdf_width=160, outpath = "output/maquis/", legend_position = "bottom")

#Test whether spatial trend is different from zero
print("Test whether spatial trend is different from zero")
mm_distance<- emtrends(glm.Meles.meles, specs = "Subunit", var = "Distance_rescaled", type = "response")
print(mm_distance)
test_results_distance <- test(mm_distance, null = 0, adjust = "fdr")
print(test_results_distance)
```

### Sus scrofa

```{r community analysis (subunit interactions) - per species models Sus scrofa}
####Sus scrofa####

Sus.scrofa <- spp_no_rare[,"Sus.scrofa"]
glm.Sus.scrofa <- glm.nb(formula = Sus.scrofa ~ Subunit * Distance_rescaled + Subunit * rescaled_Time.Diff + cosinus_Monitoring.Time.Diff, data = env_data)
coef(glm.Sus.scrofa)
coef(mva_m1)

plot_model_interaction(P.anal = env_data, m = glm.Sus.scrofa, eff2plot = "Distance_rescaled", modvar2plot = "Subunit", plot_points=FALSE, plot_residuals=FALSE, export_plot=TRUE, ylabel = NULL, fontname = "Almoni ML v5 AAA",fontsize=22, pdf_width=160, outpath = "output/maquis/", legend_position = "bottom")

sus_distance <- interact_plot(model = glm.Sus.scrofa, data=env_data, pred = Distance_rescaled, modx = Subunit, main.title = "", interval = T,plot.points = F, jitter = c(0.1,0), point.size = 3, cat.pred.point.size = 4, partial.residuals = F, colors = 'Dark2', modxvals = NULL, line.colors = 'black', point.alpha = 0.25)
sus_distance

# as.data.table(sus_distance$data)[Subunit=="Judea",Sus.scrofa]
# 1 - 0.0005676190/0.1249574967

plot_model_interaction(P.anal = env_data, m = glm.Sus.scrofa, eff2plot = "rescaled_Time.Diff", modvar2plot = "Subunit", plot_points=FALSE, plot_residuals=FALSE, export_plot=TRUE, ylabel = NULL, fontname = "Almoni ML v5 AAA",fontsize=22, pdf_width=160, outpath = "output/maquis/", legend_position = "bottom")

interact_plot(model = glm.Sus.scrofa, data=env_data, pred = rescaled_Time.Diff, modx = Subunit, main.title = "", interval = T,plot.points = F, jitter = c(0.1,0), point.size = 3, cat.pred.point.size = 4, partial.residuals = F, colors = 'Dark2', modxvals = NULL, line.colors = 'black', point.alpha = 0.25)

# generate trend of Judea on separate axes
sus_judea_plot <- plot_model_effect(P.anal = env_data, m = glm.Sus.scrofa, eff2plot = "rescaled_Time.Diff", plot_points = FALSE, plot_residuals = FALSE, export_plot = TRUE,
                                    outpath = "output/maquis/", fontname = "Almoni ML v5 AAA", at_list = list(Subunit="Judea"))
file.remove("output/maquis/mammals__Sus.scrofa_rescaled_Time.Diff.pdf")
# sus_judea_plot$layers[[2]]$aes_params$fill <- "#e3e2f0"
# sus_judea_plot$layers[[2]]$aes_params$alpha <- 0.8
sus_judea_plot$layers[[1]]$aes_params$linetype <- 2
sus_judea_plot$layers[[1]]$aes_params$colour <- "#7570b3"
sus_judea_plot$layers[[1]]$aes_params$linewidth <- 1.5
Cairo(file = paste0("output/maquis/","mammals__Sus.scrofa_rescaled_Time.Diff_Judea_only.pdf"), width = 160, height = 160*2/3, type = "PDF", units = "mm")
print(sus_judea_plot)
dev.off()

#Test whether spatial trend is different from zero
print("Test whether spatial trend is different from zero")
mm_rescaled_distance <- emtrends(glm.Sus.scrofa, specs = "Subunit", var = "Distance_rescaled", type = "response")
print(mm_rescaled_distance)
test_results_mm_distance <- test(mm_rescaled_distance, null = 0, adjust = "fdr")
print(test_results_mm_distance)

#Test whether temporal trend is different from zero
print("Test whether temporal trend is different from zero")
mm_rescaled_time <- emtrends(glm.Sus.scrofa, specs = "Subunit", var = "rescaled_Time.Diff", type = "response")
print(mm_rescaled_time)
test_results_mm_rescaled_time <- test(mm_rescaled_time, null = 0, adjust = "fdr")
print(test_results_mm_rescaled_time)
```

### Vulpes vulpes

```{r community analysis (subunit interactions) - per species models Vulpes vulpes}
#####Vulpes vulpes####

Vulpes.vulpes <- spp_no_rare[,"Vulpes.vulpes"]
glm.Vulpes.vulpes <- glm.nb(formula = Vulpes.vulpes ~ Subunit * Distance_rescaled + Subunit * rescaled_Time.Diff + cosinus_Monitoring.Time.Diff, data = env_data)
coef(glm.Vulpes.vulpes)
coef(mva_m1)

plot_model_interaction(P.anal = env_data, m = glm.Vulpes.vulpes, eff2plot = "Distance_rescaled", modvar2plot = "Subunit", plot_points=FALSE, plot_residuals=FALSE, export_plot=TRUE, ylabel = NULL, fontname = "Almoni ML v5 AAA",fontsize=22, pdf_width=160, outpath = "output/maquis/", legend_position = "bottom")

#Test whether spatial trend is different from zero
print("Test whether spatial trend is different from zero")
mm_rescaled_distance <- emtrends(glm.Vulpes.vulpes, specs = "Subunit", var = "Distance_rescaled", type = "response")
print(mm_rescaled_distance)
test_results_mm_distance <- test(mm_rescaled_distance, null = 0, adjust = "fdr")
print(test_results_mm_distance)


```

n = 1117 (total number of cameras)

total abundance = 

abundance of common species (8 species) = 22,748

### Session information

```{r}
session_info()
```

